---
output:
  github_document:
    html_preview: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```



# exuber : Econometric Analysis of Explosive Time Series

[![Build Status](https://travis-ci.org/kvasilopoulos/exuber.svg?branch=master)](https://travis-ci.org/kvasilopoulos/exuber)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/kvasilopoulos/exuber?branch=master&svg=true)](https://ci.appveyor.com/project/kvasilopoulos/exuber)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/exuber)](https://cran.r-project.org/package=exuber)
 [![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![codecov](https://codecov.io/gh/kvasilopoulos/exuber/branch/master/graph/badge.svg)](https://codecov.io/gh/kvasilopoulos/exuber)

## Description 

Testing for and dating periods of explosive dynamics (exuberance) in time series 
using recursive unit root tests as proposed by [Phillips, P. C., Shi, S. and Yu, J. (2015a)](https://doi.org/10.1111/iere.12132). Simulate a variety of periodically-collapsing 
bubble models. The estimation and simulation utilize the matrix inversion lemma 
from the recursive least squares algorithm, which results in a significant speed improvement.

## Overview

### Estimation
* `radf()` : Recursive Augmented Dickey-Fuller test
* `mc_cv()` : Monte Carlo Critical Values
* `wb_cv()` : Wild Bootstrap Critical values

### Simulation
* `sim_dgp1()` : Simulation of a single-bubble process
* `sim_dgp2()` : Simulation of a two-bubble process
* `sim_blan()` : Simulation of a Blanchard (1979) bubble process
* `sim_evans()` : Simulation of a Evans (1991) bubble process
* `sim_div()` : Simulation of dividends

## Installation

You can install the released version of exuber from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("exuber")
```

And the development version from [GitHub](https://github.com/) with:

```{r installation, eval = FALSE}
if(!require(devtools)) install.packages("devtools")
devtools::install_github("kvasilopoulos/exuber")
```
Note that development version requires compilation, so to install you will need the appropriate development tools.

* Window Users should install [Rtools](https://cran.r-project.org/bin/windows/Rtools/)
* Mac User should install [Clang or GNU Fortran](https://cran.r-project.org/bin/macosx/tools/)

If you encounter a clear bug, please file a reproducible example on [GitHub](https://github.com/kvasilopoulos/exuber/issues).

## Usage

This is a basic example which shows you how to use exuber:

```{r example, echo = TRUE, eval = FALSE}
library(exuber)
# Simulate data witn n = 100 observations
set.seed(123)
a1 <- sim_dgp1(n = 100) # one bubble
a2 <- sim_dgp2(n = 100) # two bubbles
a3 <- sim_blan(n = 100) # blanchard model
a4 <- sim_evans(n = 100) # evans model

dta <- data.frame("oneb" = a1, "twob" = a2, "blan" = a3, "evans" = a4)

ts <- radf(dta, lag = 1)

# Critical Values mc = Monte Carlo, wb= Wild Bootstrapped
# Use 500 repetions(boostraps) for faster computation
mc <- mc_cv(n = NROW(dta), nrep = 500, parallel = T)
wb <- wb_cv(dta, nboot = 500, parallel = T)

```

### Report

Report summary statistics, diagnostics and date stamping periods of mildly explosive behaviour.

```{r Report, eval = FALSE}
report(ts, mc)
diagnostics(ts, mc)
datestamp(ts, mc)

```

### Plotting

The output of plot is a list,

```{r plotting, echo=TRUE, eval = FALSE}
# All together
plot(ts, mc, plot_type = "single", breaks_x = 20)

# Individually
plot(ts, mc, plot_type = "multiple", breaks_x = 20)

library(gridExtra)
p1 <- plot(ts, mc, plot_type = "multiple", breaks_x = 20, breaks_y = 3)
do.call("grid.arrange", c(p1, ncol = 2))
```

#### License

This package is free and open source software, licensed under [GPL-3](https://github.com/kvasilopoulos/exuber/blob/master/LICENSE).

#### Code of Conduct

Please note that this project is released with a [Contributor Code of Conduct](https://github.com/kvasilopoulos/exuber/blob/master/CONDUCT.md). 
By participating in this project you agree to abide by its terms.
