---
title: "Alternative Tests: lbi_test(), ssu_test(), quantile_test()"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Alternative Tests: lbi_test(), ssu_test(), quantile_test()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, warning = FALSE, message = FALSE,
  comment = "#>"
)
```

```{r load}
library(exuber)
```

## Why not just use `radf()`

`radf()`'s GSADF statistic tests one specific alternative: a fixed explosive
AR(1) root. `lbi_test()`, `ssu_test()` and `quantile_test()` are standalone
hypothesis tests -- not built on `radf()`'s recursive core, and not fit into
the `summary()`/`datestamp()`/`tidy()`/`autoplot()` pipeline (see
`vignette("naming-and-analysis")`) -- each targeting a *different*
alternative where GSADF-style tests can lose power:

| Function | Paper | Alternative it targets |
|---|---|---|
| `lbi_test()` | Breitung & Diegel (2025) | A fixed explosive root, tested with the locally-best-invariant statistic for that specific alternative (so it can out-power GSADF exactly there). |
| `ssu_test()` | Kurozumi & Nishi (2025) | A **stochastically varying** explosive coefficient -- the root itself has a random component, not a fixed value. |
| `quantile_test()` | Wu, Shi & Wu (2025) | Explosiveness in the `tau`-th conditional quantile of `y_t` on `y_{t-1}`, not the conditional mean. |

## Same alternative, `lbi_test()` detects

```{r lbi}
y <- sim_psy1(n = 60, te = 1, tf = 60, c = 0.03, alpha = 0, seed = 1) # fixed rho = 1.03 throughout
lbi_test(y)
```

## A different alternative: where `lbi_test()` misses and `ssu_test()` doesn't

`ssu_test()` is designed for a root that itself varies stochastically over
time, not a fixed one:

`sim_psy1()`'s `coef_noise`/`coef_a` arguments generate exactly that
alternative, `rho_t = 1 + c/n + coef_a * u_t / sqrt(n)`, a random root rather
than a fixed one:

```{r stochastic-dgp}
y <- sim_psy1(n = 150, te = 75, tf = 150, c = 3, alpha = 1, seed = 2001,
              coef_noise = rnorm(149), coef_a = 4)
```

```{r ssu-vs-lbi}
ssu_test(y, level = 0.95)
lbi_test(y)
```

`ssu_test()` detects it; `lbi_test()`, built for a fixed root, does not on
this same draw. This isn't a defect in `lbi_test()` -- it's the whole reason
two tests exist: each is the (locally) most powerful test for its own
alternative, and neither dominates the other everywhere.

## Testing the quantile, not the mean

`quantile_test()` picks (or is given) a quantile `tau` and tests for
explosiveness there instead of in the conditional mean -- the setting where
that pays off is heavy-tailed innovations, where the conditional-mean
regression is least reliable, so the DGP here drives the PSY bubble with
`t(3)` shocks:

```{r quantile-test}
y_t3 <- sim_psy1(n = 100, seed = 1, e = sim_innov(99, dist = "t", df = 3))
quantile_test(y_t3, nrep = 100, seed = 1)
```

`tau = "optimal"` (the default) searches `tau_grid` and reports the quantile
with the strongest signal, shown here fixed at a specific value for a
faster, reproducible example.

## Which to reach for

- A genuinely fixed explosive root, and want a test with power advantages
  over standard GSADF for exactly that case: `lbi_test()`.
- Suspect the explosive root itself is noisy/time-varying rather than
  constant: `ssu_test()`.
- Suspect explosiveness shows up more in the tails (or a specific quantile)
  of the distribution than in the mean: `quantile_test()`.
- Not sure which alternative applies, or want the most widely used
  benchmark: `radf()`'s GSADF remains the default first test to run.
