---
title: "Alternative Tests: lbi_test(), ssu_test(), quantile_test()"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Alternative Tests: lbi_test(), ssu_test(), quantile_test()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, warning = FALSE, message = FALSE,
  comment = "#>"
)
```

```{r load}
library(exuber)
```

## Why not just use `radf()`

`radf()`'s GSADF statistic tests one specific alternative: a fixed explosive
AR(1) root. `lbi_test()`, `ssu_test()` and `quantile_test()` are standalone
hypothesis tests -- not built on `radf()`'s recursive core, and not fit into
the `summary()`/`datestamp()`/`tidy()`/`autoplot()` pipeline (see
`vignette("naming-and-analysis")`) -- each targeting a *different*
alternative where GSADF-style tests can lose power:

| Function | Paper | Alternative it targets |
|---|---|---|
| `lbi_test()` | Breitung & Diegel (2025) | A fixed explosive root, tested with the locally-best-invariant statistic for that specific alternative (so it can out-power GSADF exactly there). |
| `ssu_test()` | Kurozumi & Nishi (2025) | A **stochastically varying** explosive coefficient -- the root itself has a random component, not a fixed value. |
| `quantile_test()` | Wu, Shi & Wu (2025) | Explosiveness in the `tau`-th conditional quantile of `y_t` on `y_{t-1}`, not the conditional mean. |

## Same alternative, `lbi_test()` detects

```{r lbi}
set.seed(1)
n <- 60
y <- 100 * 1.03^(1:n) + cumsum(rnorm(n, sd = 1)) # fixed rho = 1.03
lbi_test(y)
```

## A different alternative: where `lbi_test()` misses and `ssu_test()` doesn't

`ssu_test()` is designed for a root that itself varies stochastically over
time, not a fixed one:

```{r stochastic-dgp}
make_stochastic_bubble <- function(n, te_frac = 0.5, c1 = 3, a = 4) {
  y <- numeric(n)
  y[1] <- rnorm(1)
  Te <- round(te_frac * n)
  for (t in 2:n) {
    if (t <= Te) {
      y[t] <- y[t - 1] + rnorm(1)
    } else {
      rho_t <- 1 + c1 / n + a * rnorm(1) / sqrt(n) # random root, not fixed
      y[t] <- rho_t * y[t - 1] + rnorm(1)
    }
  }
  y
}
set.seed(2001)
y <- make_stochastic_bubble(150)
```

```{r ssu-vs-lbi}
ssu_test(y, level = 0.95)
lbi_test(y)
```

`ssu_test()` detects it; `lbi_test()`, built for a fixed root, does not on
this same draw. This isn't a defect in `lbi_test()` -- it's the whole reason
two tests exist: each is the (locally) most powerful test for its own
alternative, and neither dominates the other everywhere.

## Testing the quantile, not the mean

`quantile_test()` picks (or is given) a quantile `tau` and tests for
explosiveness there instead of in the conditional mean:

```{r quantile-test}
quantile_test(sim_data$psy2, nrep = 100, seed = 1)
```

`tau = "optimal"` (the default) searches `tau_grid` and reports the quantile
with the strongest signal, shown here fixed at a specific value for a
faster, reproducible example.

## Which to reach for

- A genuinely fixed explosive root, and want a test with power advantages
  over standard GSADF for exactly that case: `lbi_test()`.
- Suspect the explosive root itself is noisy/time-varying rather than
  constant: `ssu_test()`.
- Suspect explosiveness shows up more in the tails (or a specific quantile)
  of the distribution than in the mean: `quantile_test()`.
- Not sure which alternative applies, or want the most widely used
  benchmark: `radf()`'s GSADF remains the default first test to run.
