---
title: "Bivariate Bubble Relationships: cobubble_test() and contagion_reg()"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bivariate Bubble Relationships: cobubble_test() and contagion_reg()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, warning = FALSE, message = FALSE,
  comment = "#>"
)
```

```{r load}
library(exuber)
```

## Two different questions about two series

Both functions here look at a *pair* of series that each contain (or might
contain) an explosive episode, but they ask different questions and answer
them differently:

- `cobubble_test()` (Evripidou, Harvey, Leybourne & Sollis 2022) is a formal
  **hypothesis test**: are `y` and `x`'s explosive episodes the *same*
  episode, i.e. is `y_t - alpha - beta * x_{t-i}` stationary for some
  lag/lead `i`? This is a KPSS-type test -- the null hypothesis is
  co-explosivity (stationary residual), so *rejecting* means the two series'
  bubbles are not the same underlying process.
- `contagion_reg()` (Greenaway-McGrevy & Phillips 2016) performs no formal
  inference at all. It estimates a **time-varying** contagion coefficient
  `delta_2(r)`: at each point `r` in the sample, how strongly a "peripheral"
  series `y`'s (fixed-window) AR(1) coefficient co-moves with a "core"
  series' own coefficient, via a Nadaraya-Watson kernel regression.

Neither carries the `radf_obj` class -- see `vignette("naming-and-analysis")`.

## `cobubble_test()`: are these the same bubble?

`sim_data$psy1` and `sim_data$psy2` are independently simulated explosive
episodes, so a correct test should reject co-explosivity between them:

```{r cobubble}
res <- cobubble_test(sim_data$psy1, sim_data$psy2, nboot = 199, seed = 1)
res
```

It does: `S` comfortably exceeds its (wild-bootstrap, heteroskedasticity-robust)
critical value, and co-explosivity is rejected -- correctly, since these two
series share no common bubble process by construction.

## `contagion_reg()`: how strongly do they co-move, and when?

```{r contagion}
cr <- contagion_reg(sim_data$psy1, sim_data$psy2, d = 0)
cr
```

`cr$delta2` is the full estimated path over `cr$r_grid`, not just the range
`print()` shows:

```{r contagion-plot, fig.width = 6, fig.height = 4}
plot(cr$r_grid, cr$delta2, type = "l",
     xlab = "r (fraction of sample)", ylab = expression(delta[2](r)),
     main = "Estimated time-varying contagion coefficient")
```

For contrast, a series with no relationship to `psy2` at all -- a plain
random walk -- gives a visibly different, wider-ranging path that also
crosses zero:

```{r contagion-null}
set.seed(5)
indep <- cumsum(rnorm(100))
cr_null <- contagion_reg(indep, sim_data$psy2, d = 0)
range(cr_null$delta2)
range(cr$delta2)
```

## Which to reach for

- Want a yes/no answer, with a critical value, to "do these two series share
  the same explosive episode": `cobubble_test()`.
- Want to see *how* the strength of comovement between two series' AR
  coefficients evolves over the sample (e.g. to visualize contagion
  building up before a joint collapse), with no formal test attached:
  `contagion_reg()`.
