---
title: "Bivariate Bubble Relationships: cobubble_test() and contagion_reg()"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bivariate Bubble Relationships: cobubble_test() and contagion_reg()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, warning = FALSE, message = FALSE,
  comment = "#>"
)
```

```{r load}
library(exuber)
```

## Two different questions about two series

Both functions look at a pair of series that each contain, or might contain, an explosive episode. They ask different questions.

- `cobubble_test()` (Evripidou, Harvey, Leybourne & Sollis 2022) is a formal hypothesis test of whether the explosive episodes of `y` and `x` are the same episode, that is, whether `y_t - alpha - beta * x_{t-i}` is stationary for some lag or lead `i`. It is a KPSS-type test whose null hypothesis is co-explosivity (a stationary residual). Rejecting the null therefore means that the bubbles in the two series do not come from the same underlying process.
- `contagion_reg()` (Greenaway-McGrevy & Phillips 2016) does no formal inference. It estimates a time-varying contagion coefficient `delta_2(r)` with a Nadaraya-Watson kernel regression. At each point `r` in the sample, the coefficient measures how strongly the fixed-window AR(1) coefficient of a "peripheral" series `y` moves with the coefficient of a "core" series.

Neither function returns the `radf_obj` class (see `vignette("naming-and-analysis")`).

## `cobubble_test()`: are these the same bubble?

`sim_coexplosive()` implements the data generating process of Evripidou et al.: `y` is a linear function of an explosive `x` plus noise. The pair is co-explosive, so a correct test should not reject:

```{r cobubble}
xy <- sim_coexplosive(n = 100, seed = 123)
res <- cobubble_test(xy$y, xy$x, nboot = 199, seed = 1)
res
```

For contrast, `sim_data$psy1` and `sim_data$psy2` are simulated independently, so by construction they share no bubble process:

```{r cobubble-indep}
cobubble_test(sim_data$psy1, sim_data$psy2, nboot = 199, seed = 1)
```

Here `S` clearly exceeds its wild-bootstrap critical value, which is robust to heteroskedasticity, and co-explosivity is rejected, as it should be.

## `contagion_reg()`: how strongly do they co-move, and when?

For the co-explosive pair, the AR(1) coefficient of `y` follows that of `x` almost one for one throughout the sample:

```{r contagion}
cr <- contagion_reg(xy$y, xy$x, d = 0)
cr
```

`cr$delta2` holds the whole estimated path over `cr$r_grid`, while `print()` shows only its range:

```{r contagion-plot, fig.width = 6, fig.height = 4}
plot(cr$r_grid, cr$delta2, type = "l",
     xlab = "r (fraction of sample)", ylab = expression(delta[2](r)),
     main = "Estimated time-varying contagion coefficient")
```

For contrast, the two independent `sim_data` series give a coefficient path that stays far below one:

```{r contagion-null}
cr_null <- contagion_reg(sim_data$psy1, sim_data$psy2, d = 0)
range(cr_null$delta2)
range(cr$delta2)
```

## Which to reach for

- If you want a yes or no answer, with a critical value, to the question of whether two series share the same explosive episode, use `cobubble_test()`.
- If you want to see how the comovement between the AR coefficients of two series changes over the sample, for example to watch contagion build up before a joint collapse, and you do not need a formal test, use `contagion_reg()`.
