---
title: "Dating Methods: Alternatives to datestamp()"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Dating Methods: Alternatives to datestamp()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, warning = FALSE, message = FALSE,
  comment = "#>"
)
```

```{r load}
library(exuber)
```

## Why these exist alongside `datestamp()`

`datestamp()` applies PSY's own rule to a `radf()` result: a bubble runs from
the first point the recursive statistic crosses its critical value to the
first point it drops back below. That rule is simple and well understood,
but it is not the only way to *date* a bubble once you already believe one is
there. The `dating_*()` family instead fits an explicit regime model --
unit-root, then explosive, then unit-root again -- directly to the raw
series by minimizing residual sum of squares (SSR), and picks the break
dates that fit that model best. They take no critical value at all: given a
window believed to contain (at most) one bubble, they answer "where exactly
does it start and end," not "is there one."

| Function | Paper | Idea |
|---|---|---|
| `dating_hls()` | Harvey, Leybourne & Sollis (2017) | Fits 4 candidate regime-dummy models (with/without a distinct collapse regime, with/without recovery) via closed-form segment SSR; BIC picks among them. |
| `dating_knp()` | Kejriwal, Nguyen & Perron (2025) | Same model as HLS's Model 2, but proves the plain SSR minimizer is *inconsistent* -- it converges to the collapse date, not the origination date -- and fixes it by omitting one squared residual from the objective. |
| `dating_pdc()` | Pang, Du & Chong (2021); Kurozumi & Skrobotov (2023) | Assumes a fixed 3- or 4-regime structure and finds each breakpoint *sequentially* in closed form (collapse first, since it is stochastically dominant), with no BIC step. |
| `dating_hlw()` | Harvey, Leybourne & Whitehouse (2020) | A wrapper: runs `radf()`/`datestamp()` first to find *how many* episodes and roughly where, then applies HLS-style fitting independently within each detected window. |

All four take no `radf_cv`, so they don't plug into `summary()`/`tidy()`/
`autoplot()` -- see `vignette("naming-and-analysis")` for the full pipeline
picture. Each prints its own dating table instead.

## A single bubble, four verdicts

One simulated series: a unit-root run-up, a genuine explosive regime
(`rho = 1.03`, true origination at 51), then a volatile collapse back to a
unit root (true collapse at 80).

```{r dgp}
set.seed(11)
n1 <- 50; n2 <- 30; n3 <- 20
expansion <- 100 * 1.03^(1:n2) + cumsum(rnorm(n2, sd = 1))
y <- c(cumsum(rnorm(n1)), expansion, expansion[n2] + cumsum(rnorm(n3, sd = 1.5)))
```

True origination is 51, true collapse is 80. Running all four:

```{r hls}
dating_hls(y, trim = 0.05)
```

```{r knp}
dating_knp(y, trim = 0.05)
```

```{r pdc}
dating_pdc(y, regimes = 3, trim = 0.05)
```

```{r hlw}
dating_hlw(y, trim = 0.1, nboot = 199, seed = 1)
```

On this draw, `dating_pdc()` lands closest to both true dates (50, 79).
`dating_hls()` gets the origination right (50) but its BIC-selected model
puts the collapse far too early (55) -- a real, visible failure mode, not
a contrived one. `dating_hlw()`'s preliminary `datestamp()` step splits the
episode into two separate windows here, and its second window's collapse
(80) is exact even though the first window is spurious. `dating_knp()`'s
origination (28) is the least accurate of the four on this particular draw.

The point of running all four side by side isn't that one is "correct" --
it's that these are genuinely different estimators with different failure
modes, and disagreement between them on real data is informative, not a bug
to be resolved by picking a favorite.

## Which to reach for

- Believe there is exactly **one** bubble in the window and want the
  best-fitting regime model (with or without a distinct collapse/recovery
  regime): `dating_hls()`.
- Same setting, but origination-date accuracy matters more than
  collapse-date accuracy (KNP's own finding: the naive estimator is biased
  toward the collapse date): `dating_knp()`.
- Want closed-form, no BIC model search, and are willing to fix the number
  of regimes up front: `dating_pdc()` (add `weights` for the
  volatility-corrected variant).
- Don't know how many episodes there are, or want dating anchored to an
  actual `radf()`/`datestamp()` detection first: `dating_hlw()`.
