---
title: "Dating Methods: Alternatives to datestamp()"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Dating Methods: Alternatives to datestamp()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, warning = FALSE, message = FALSE,
  comment = "#>"
)
```

```{r load}
library(exuber)
```

## Why these exist alongside `datestamp()`

`datestamp()` applies PSY's own rule to a `radf()` result: a bubble runs from
the first point the recursive statistic crosses its critical value to the
first point it drops back below. That rule is simple and well understood,
but it is not the only way to *date* a bubble once you already believe one is
there. The `dating_*()` family instead fits an explicit regime model --
unit-root, then explosive, then unit-root again -- directly to the raw
series by minimizing residual sum of squares (SSR), and picks the break
dates that fit that model best. They take no critical value at all: given a
window believed to contain (at most) one bubble, they answer "where exactly
does it start and end," not "is there one."

| Function | Paper | Idea |
|---|---|---|
| `dating_hls()` | Harvey, Leybourne & Sollis (2017) | Fits 4 candidate regime-dummy models (with/without a distinct collapse regime, with/without recovery) via closed-form segment SSR; BIC picks among them. |
| `dating_knp()` | Kejriwal, Nguyen & Perron (2025) | Same model as HLS's Model 2, but proves the plain SSR minimizer is *inconsistent* -- it converges to the collapse date, not the origination date -- and fixes it by omitting one squared residual from the objective. |
| `dating_pdc()` | Pang, Du & Chong (2021); Kurozumi & Skrobotov (2023) | Assumes a fixed 3- or 4-regime structure and finds each breakpoint *sequentially* in closed form (collapse first, since it is stochastically dominant), with no BIC step. |
| `dating_hlw()` | Harvey, Leybourne & Whitehouse (2020) | A wrapper: runs `radf()`/`datestamp()` first to find *how many* episodes and roughly where, then applies HLS-style fitting independently within each detected window. |

All four take no `radf_cv`, so they don't plug into `summary()`/`tidy()`/
`autoplot()` -- see `vignette("naming-and-analysis")` for the full pipeline
picture. Each prints its own dating table instead.

## A single bubble, four verdicts

One simulated series from `sim_ps1()`, Phillips & Shi (2018)'s own
single-bubble DGP: a unit-root run-up, a genuine explosive regime (true
origination at 40), a mildly-integrated collapse regime (starting at 61) and
a return to a unit root (recovery at 70) -- exactly the regime structure
these estimators are built for.

```{r dgp}
y <- sim_ps1(n = 100, seed = 1)
```

True origination is 40, true collapse is 61. Running all four:

```{r hls}
dating_hls(y, trim = 0.05)
```

```{r knp}
dating_knp(y, trim = 0.05)
```

```{r pdc}
dating_pdc(y, regimes = 3, trim = 0.05)
```

```{r hlw}
dating_hlw(y, trim = 0.1, nboot = 199, seed = 1)
```

On this draw, `dating_hls()` and `dating_hlw()` both select the 4-regime
model and land within a point of every true date (39/60/70 against
40/61/70); `dating_pdc()` is a point or two early on both (38/59).
`dating_knp()` is the visible failure mode here: its model assumes an
*instantaneous* collapse (unit root resuming from a shifted level), so
`sim_ps1()`'s ten-period mildly-integrated crash (61-70) is a regime it
has no room for, and it dates that crash as the episode instead,
reporting 60/70 -- a real, not contrived, consequence of a model
mismatch, not of the estimator's bias correction.

The point of running all four side by side isn't that one is "correct" --
it's that these are genuinely different estimators with different failure
modes, and disagreement between them on real data is informative, not a bug
to be resolved by picking a favorite.

## Which to reach for

- Believe there is exactly **one** bubble in the window and want the
  best-fitting regime model (with or without a distinct collapse/recovery
  regime): `dating_hls()`.
- Same setting, but origination-date accuracy matters more than
  collapse-date accuracy (KNP's own finding: the naive estimator is biased
  toward the collapse date): `dating_knp()`.
- Want closed-form, no BIC model search, and are willing to fix the number
  of regimes up front: `dating_pdc()` (add `weights` for the
  volatility-corrected variant).
- Don't know how many episodes there are, or want dating anchored to an
  actual `radf()`/`datestamp()` detection first: `dating_hlw()`.
