---
title: "Experimental Methods: radf_recovery() and datestamp(option = 'svadf')"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Experimental Methods: radf_recovery() and datestamp(option = 'svadf')}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, warning = FALSE, message = FALSE,
  comment = "#>"
)
```

```{r load}
library(exuber)
```

## Why "experimental" is a real, load-bearing label here

Most of exuber's methods implement a peer-reviewed paper's own procedure and
pass the package's standard validation pass (formula-exact check against a
brute-force reimplementation, published-table lookup, Monte Carlo size,
power against a genuine alternative). `radf_recovery()` and
`datestamp(option = "svadf")` both run that same pass and both come back
genuinely useful -- but each has
one specific, disclosed gap that keeps it below that bar. They print
`r lifecycle::badge("experimental")` and emit a caveat message at call time
for exactly that reason; treat their output as directional, not as
calibrated as the rest of the package.

## `radf_recovery()`: dating a collapse *and* a recovery

Phillips & Shi (2014)'s reverse-regression idea: reverse the series in time,
run the same BSADF recursion `radf()` already computes, and map the
crossing dates back. A collapse-then-recovery episode, reversed, turns the
collapse into an explosive regime and the recovery into its unwind -- so the
existing forward machinery, run backwards, dates both.

```{r recovery}
set.seed(2)
n1 <- 40; n2 <- 25; n3 <- 35
expansion <- 100 * 1.03^(1:n1) + cumsum(rnorm(n1, sd = 1))
collapse <- expansion[n1] * 0.5^((1:n2) / n2) + cumsum(rnorm(n2, sd = 1))
recovery <- collapse[n2] + cumsum(rnorm(n3, sd = 1)) + (1:n3) * 0.5
y <- c(expansion, collapse, recovery) # expansion -> collapse -> recovery
res <- radf_recovery(y, minw = 15, nrep = 200, seed = 1)
res
```

`f_c` (crisis onset) lands near the true collapse start (40) and `f_r`
(recovery) after it (62), in the right order -- by construction, since the
down-crossing search only ever starts at the up-crossing. The disclosed gap:
`f_c` and the overall false-detection rate are exploratory pending further
validation (see `?radf_recovery`, Caveats section) -- the date-ordering
property is solid, the false-alarm calibration is not yet.

## `datestamp(option = "svadf")`: a non-peer-reviewed preprint

Sarkar & Wells (2026, arXiv, not yet peer reviewed) -- flagged explicitly
because that is a different evidentiary bar than every other paper this
package implements. Its statistic turns out to be exactly `radf()`'s own
`badf` sequence compared against two closed-form, sample-size-only
thresholds from the paper's own applied methodology (no new estimation
machinery needed), which is why it was cheap to add -- as a `datestamp()`
option rather than a separate entry point, despite the preprint caveat.

```{r svadf}
res <- radf(sim_data, lag = 0)
datestamp(res, option = "svadf", min_duration = psy_ds(nrow(sim_data)))
```

`psy1` and `psy2` get clean origination/collapse dates; `evans`, `div` and
`blan` don't cross the threshold on this panel at all -- again a realistic,
not cherry-picked, mixed result.

## Using them responsibly

Both are worth using -- `radf_recovery()`'s date *ordering* result and
`datestamp(option = "svadf")`'s point statistic are both solid -- but
neither should be the sole basis for a claim about false-alarm rates or
exact calibration.
Prefer `radf()`/`datestamp()` or one of the peer-reviewed alternatives in
`vignette("alternative-tests")`/`vignette("dating-methods")` when that
matters, and treat these two as a second opinion rather than the primary
one until their own caveats are resolved (tracked in
`docs/enhancements/dating-and-root-inference.md` and
`volatility-robustness.md`).
