---
title: "Intro to exuber"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Intro to exuber}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message=FALSE
)
options(cli.width = 80)
```

```{r echo=FALSE, message=FALSE}
library(exuber)
```

This vignette uses the `datasets::EuStockMarkets` data, which contain the daily closing prices of four major European stock indices: the German DAX, the Swiss SMI, the French CAC and the UK FTSE (see `?EuStockMarkets`). The data are sampled in business time, so weekends and holidays are omitted. We want weekly observations, so we aggregate to a weekly frequency, which reduces the sample from 1860 to 372 observations.

```{r dataset}
stocks <- aggregate(EuStockMarkets, nfrequency = 52, mean)
```

## Estimation

We estimate the recursive augmented Dickey-Fuller test on these series with one lag.

```{r estimation}
est_stocks <- radf(stocks, lag = 1)
```

## Analysis

`summary()` prints the test statistics together with the critical values at the 10%, 5% and 1% significance levels. A critical value is the threshold a statistic must exceed before we reject the null hypothesis of a unit root in favor of explosive behavior. When `cv` is omitted, `summary()`, `diagnostics()`, `datestamp()` and `autoplot()` fetch precomputed Monte Carlo critical values for the relevant `(n, lag)` from a shared store. The store covers lags 0 to 4 and samples of up to 4000 observations. The first fetch needs network access, and later calls read from the disk cache. Here we simulate the critical values locally with `radf_mc_cv()` and pass the result through `cv` to every downstream call. This offline route is also the one to use for other lags or larger samples.

```{r cv}
cv_stocks <- radf_mc_cv(NROW(stocks), lag = 1)
```

```{r summary}
summary(est_stocks, cv = cv_stocks)
```

All four stocks appear to show exuberant behavior, and `diagnostics()` confirms this by reporting which series reject the null. It is most useful when there are many series.

```{r diagnostics}
diagnostics(est_stocks, cv = cv_stocks)
```


To find out when the exuberance occurred, we use `datestamp()`, which takes the same arguments as `summary()` and `diagnostics()`.

```{r datestamp}
# Minimum duration of an explosive period
rot = psy_ds(stocks) # log(n) ~ rule of thumb

dstamp_stocks <- datestamp(est_stocks, cv = cv_stocks, min_duration = rot)
dstamp_stocks
```
We can also extract the datestamp as a dummy variable, where 1 marks exuberance and 0 marks its absence.

```{r datestamp-dummy}
dummy <- attr(dstamp_stocks, "dummy")
tail(dummy)
```

`datestamp()` is not the only way to obtain dates, but the other two approaches answer different questions rather than the same one. The `dating_*()` family (`vignette("dating-methods")`) fits an explicit regime model to date a bubble that you already believe is there, and it does not test whether one exists. The `monitor()` and `monitor_*()` family (`vignette("monitoring")`) detects bubbles in real time by watching new observations one at a time, instead of dating a finished sample. `vignette("naming-and-analysis")` shows how every function in the package relates to `datestamp()`.

## Plotting

`autoplot()` returns a faceted ggplot2 object for all the series that reject the null hypothesis at the 5% significance level.

```{r plot-radf, fig.width = 9}
autoplot(est_stocks, cv = cv_stocks)
```

Finally, we can plot only the periods of exuberance. Plotting the datestamp object is useful when there are many series and we want to see the explosive episodes in all of them.

```{r plot-datestaemp, fig.width = 7, fig.height=2}
datestamp(est_stocks, cv = cv_stocks) %>%
  autoplot()
```
