---
title: "Intro to exuber"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(exuber)
```

We are goint to use the EuStockMarkets dataset, which contains the daily closing prices of four
major European stock indices: Germany DAX (Ibis), Switzerland SMI, France CAC, and UK FTSE.


The data are sampled in business time, i.e., weekends and holidays are omitted. In this particular
exercise we want to focus on weekly observations, thus we need to aggregate to a weekly closing prices.

```{r dataset}
dataset <- 
  EuStockMarkets %>% 
  aggregate(nfrequency = 52, mean)
```

The number of observations reduce from 1860 to 372. Next we estimate the 


```{r estimation}
est_dt <-
  dataset %>% 
  radf(lag = 1)

```

```{r summary}
summary(est_dt)
```

The `autoplot.radf` function returns a list of ggplot2, thus we need to grob the list into a
single graph. Function `ggarrange

```{r datestamp, fig.width = 9}
est_dt %>% 
  autoplot() %>% 
  ggarrange(ncol = 2)
```
ge(ncol = 2)
```
