---
title: "Real-Time Monitoring"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Real-Time Monitoring}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, warning = FALSE, message = FALSE,
  comment = "#>"
)
```

```{r load}
library(exuber)
```

## Monitoring vs. testing

`radf()`/`datestamp()` and the `dating_*()` family (see
`vignette("dating-methods")`) all work on a *finished* sample: they answer
"was there a bubble, and when" after every observation is already in hand.
The `monitor_*()` family answers a different, real-time question: fix a
training window `[1, T*]` believed free of exuberance, calibrate a boundary
on it, then watch each new observation `T*+1, T*+2, ...` and raise an alarm
the first time the boundary is breached. All four functions share this
`r_star`/alarm/`alarm_date` shape; they differ in what statistic they
monitor and how the boundary is calibrated.

| Function | Statistic monitored | Boundary | Static, full-sample counterpart |
|---|---|---|---|
| `monitor()` | `radf()`'s own `badf`/`bsadf` recursion | `"bootstrap"` (Phillips & Shi 2020 wild-bootstrap quantile), `"kurozumi"` (closed-form, Kurozumi 2020), or `"fluc"` (closed-form, Homm & Breitung 2012) | `radf()`/`datestamp()` -- not a `_test()`, but the same recursive-ADF core |
| `monitor_cusum()` | A CUSUM of the training-window-standardized series | Homm & Breitung (2012)'s asymptotic (or finite-sample) constant | none -- Homm & Breitung's CUSUM boundary is inherently a training/monitoring construction, with no full-sample form |
| `monitor_lbi()` | Breitung & Diegel (2025)'s locally-best-invariant CUSUM (`mCUSUM`/`wCUSUM`, via `c_bar`) | Their Table 1 constant | `lbi_test()`, the static version of the same statistic |
| `monitor_quantile()` | A recursive quantile regression at `tau` | A simulated first-crossing boundary (Wu, Shi & Wu 2025) | `quantile_test()`, the static version of the same statistic |

`monitor()` reuses `badf`/`bsadf` directly (the same recursive-ADF core as
the `radf_*()` family) but is named for what it *does*, not that internal
detail -- see `vignette("naming-and-analysis")` for why. Two of its three
siblings (`monitor_lbi()`, `monitor_quantile()`) are the sequential
extension of an existing static test of the same name minus the
`monitor_` prefix; `monitor_cusum()` has no such counterpart, since its
source paper (Homm & Breitung 2012) proposed CUSUM as a monitoring
detector only.

## The same bubble, five monitors

A training window of pure random walk (`T* = 100`), followed by more random
walk, then a genuine explosive regime (`rho = 1.04`) starting at `t = 150`:

```{r dgp}
make_bubble_series <- function(n, T_star, bstart, rho = 1.04) {
  y <- numeric(n)
  y[seq_len(T_star)] <- cumsum(rnorm(T_star))
  for (t in (T_star + 1):n) {
    y[t] <- if (t < bstart) y[t - 1] + rnorm(1) else rho * y[t - 1] + rnorm(1)
  }
  y
}
set.seed(7)
y <- make_bubble_series(200, T_star = 100, bstart = 150)
```

```{r monitors}
monitor_lbi(y, r_star = 100)
monitor_cusum(y, r_star = 0.5)
monitor_quantile(y, tau = 0.5, nrep = 200, seed = 1)
monitor(y, r_star = 0.5, nboot = 200, seed = 1)
monitor(y, r_star = 0.5, boundary = "kurozumi")
```

`monitor_lbi()` and `monitor_quantile()` each have a static, full-sample
sibling that asks the retrospective version of the same question -- run on
the whole series rather than watching for a first crossing:

```{r static-siblings}
lbi_test(y)
quantile_test(y, tau = 0.5)
```

Every monitor here alarms within about 15 points of the true bubble start
(150), none before it -- that "never before `T*` (or the true start)" property
is exactly what each function's own test suite checks under the null. Alarm
*timing* differs by design: `monitor()`'s ADF-family statistics tend to
detect mid-sample bubbles fastest (the literature's own finding, e.g.
Kurozumi 2020/2021), while CUSUM-type detectors (`monitor_cusum()`,
`monitor_lbi()`) are typically slower but computationally simpler and don't
need a bootstrap.

## Which to reach for

- Fastest detection, willing to pay for a wild bootstrap per call:
  `monitor(boundary = "bootstrap")` (the default).
- Same statistic, no bootstrap, an off-the-shelf published constant instead:
  `monitor(boundary = "kurozumi")` or `boundary = "fluc"`.
- A simpler CUSUM-based alternative with its own closed-form boundary:
  `monitor_cusum()`, or `monitor_lbi()` for Breitung & Diegel's
  locally-best-invariant version (`c_bar > 0` trades a little size for
  power against slow-building bubbles).
- Monitoring a specific quantile of the distribution rather than the mean
  behavior: `monitor_quantile()`.
