---
title: "Real-Time Monitoring"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Real-Time Monitoring}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, warning = FALSE, message = FALSE,
  comment = "#>"
)
```

```{r load}
library(exuber)
```

## Monitoring and testing

`radf()`, `datestamp()` and the `dating_*()` family (see `vignette("dating-methods")`) all work on a finished sample. They answer the question of whether there was a bubble, and when, after every observation is already in hand. The `monitor_*()` family answers a real-time question instead. We fix a training window `[1, T*]` that we believe is free of exuberance and calibrate a boundary on it. We then watch each new observation `T*+1, T*+2, ...` and raise an alarm the first time the boundary is crossed. All four functions share this `r_star`, alarm and `alarm_date` structure. They differ in the statistic they monitor and in how they calibrate the boundary.

| Function | Statistic monitored | Boundary | Static, full-sample counterpart |
|---|---|---|---|
| `monitor()` | `radf()`'s own `badf`/`bsadf` recursion | `"bootstrap"` (Phillips & Shi 2020 wild-bootstrap quantile), `"kurozumi"` (closed-form, Kurozumi 2020), or `"fluc"` (closed-form, Homm & Breitung 2012) | `radf()` and `datestamp()`. It is not a `_test()` function, but it uses the same recursive ADF core |
| `monitor_cusum()` | A CUSUM of the training-window-standardized series | Homm & Breitung (2012)'s asymptotic (or finite-sample) constant | None. The CUSUM boundary of Homm & Breitung is a training and monitoring construction by design and has no full-sample form |
| `monitor_lbi()` | Breitung & Diegel (2025)'s locally-best-invariant CUSUM (`mCUSUM`/`wCUSUM`, via `c_bar`) | Their Table 1 constant | `lbi_test()`, the static version of the same statistic |
| `monitor_quantile()` | A recursive quantile regression at `tau` | A simulated first-crossing boundary (Wu, Shi & Wu 2025) | `quantile_test()`, the static version of the same statistic |

`monitor()` reuses `badf` and `bsadf` directly, which is the same recursive ADF core as in the `radf_*()` family. It is named after what it does and not after that internal detail (see `vignette("naming-and-analysis")` for the reasoning). Two of its three siblings, `monitor_lbi()` and `monitor_quantile()`, are the sequential extension of an existing static test with the same name minus the `monitor_` prefix. `monitor_cusum()` has no such counterpart, because its source paper (Homm & Breitung 2012) proposed the CUSUM only as a monitoring detector.

## One bubble, five monitors

The series starts with a training window of pure random walk (`T* = 100`) and continues as a random walk until `t = 150`. From then until the end of the sample it follows an explosive regime (`rho = 1.04`). We generate it with `sim_psy1()`, placing the bubble after the training window and leaving out the collapse:

```{r dgp}
y <- sim_psy1(n = 200, te = 150, tf = 200, c = 0.04, alpha = 0, seed = 7)
```

```{r monitors}
monitor_lbi(y, r_star = 100)
monitor_cusum(y, r_star = 0.5)
monitor_quantile(y, tau = 0.5, nrep = 200, seed = 1)
monitor(y, r_star = 0.5, nboot = 200, seed = 1)
monitor(y, r_star = 0.5, boundary = "kurozumi")
```

`monitor_lbi()` and `monitor_quantile()` each have a static, full-sample counterpart that asks the retrospective version of the same question. It runs on the whole series and does not wait for a first crossing:

```{r static-siblings}
lbi_test(y)
quantile_test(y, tau = 0.5)
```

Every monitor alarms within about 15 observations of the true bubble start (150), and none alarms before it. Each function's own test suite checks the absence of alarms before `T*` (or the true start) under the null. The timing of the alarm differs by design. The ADF-family statistics in `monitor()` tend to detect bubbles in the middle of the sample fastest, as the literature finds (for example Kurozumi 2020, 2021). The CUSUM-type detectors (`monitor_cusum()` and `monitor_lbi()`) are typically slower, but they are computationally simpler and need no bootstrap.

## Which to reach for

- If you want the fastest detection and can afford a wild bootstrap on each call, use `monitor(boundary = "bootstrap")`, which is the default.
- If you want the same statistic without a bootstrap, using a published constant instead, use `monitor(boundary = "kurozumi")` or `monitor(boundary = "fluc")`.
- If you want a simpler CUSUM-based alternative with its own closed-form boundary, use `monitor_cusum()`. For the locally best invariant version of Breitung & Diegel, use `monitor_lbi()`, where `c_bar > 0` gives up a little size in exchange for power against slowly building bubbles.
- If you want to monitor a specific quantile of the distribution instead of the mean, use `monitor_quantile()`.
