---
title: "Naming Conventions and the Analysis/Tidying/Plotting Pipeline"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Naming Conventions and the Analysis/Tidying/Plotting Pipeline}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, warning = FALSE, message = FALSE,
  comment = "#>"
)
```

```{r load}
library(exuber)
```

## Why this exists

`exuber` started as one test (`radf()`, the recursive ADF/SADF/GSADF/BSADF
statistic of Phillips, Shi & Yu 2015) and grew, through a long research
programme, into roughly 25 functions covering a dozen papers' worth of
related-but-distinct methodology: alternative tests, dating procedures,
monitoring schemes, root inference. Every one of them used to be named
`radf_<something>()`, which was accurate for some and misleading for others --
a `radf_` prefix reads as "this is a recursive-ADF statistic," but several of
these functions are not that at all. This vignette documents the naming
scheme that replaced it, and -- more usefully -- which functions can actually
be plugged into the shared `summary()`/`datestamp()`/`tidy()`/`autoplot()`
pipeline built for `radf()`, and which have their own, differently-shaped
output instead.

## The naming scheme

| Pattern | Means | Examples |
|---|---|---|
| `radf_` prefix | Genuinely built on the recursive-ADF core: calls `radf()` directly, or reuses its `badf`/`bsadf` recursion | `radf()`, `radf_tt()`, `radf_sign()`, `radf_common()`, `radf_kp()`, `radf_recovery()`, and the `_cv`/`_mc`/`_sb`/`_wb` critical-value engines |
| `_test` suffix | A standalone hypothesis test with its own null distribution, not built on the recursive-ADF core | `lbi_test()`, `ssu_test()`, `quantile_test()`, `cobubble_test()` |
| `dating_` prefix | Point-estimation / model-selection dating, no formal hypothesis test at all | `dating_hls()`, `dating_hlw()`, `dating_knp()`, `dating_pdc()` |
| `monitor`/`monitor_` prefix | Real-time/sequential detection -- grouped by *what it does*, not by internal mechanism. `monitor()` is this family's flagship, the same role `radf()` plays for the `radf_` family: it reuses `badf`/`bsadf` directly (genuinely ADF-family internals) but carries no `radf`/`sadf` token at all, specifically so it reads as "the real-time monitor," not as a `radf_` variant -- see below | `monitor()`, `monitor_cusum()`, `monitor_lbi()`, `monitor_quantile()` |
| `root_` prefix | Confidence-interval inference on the *magnitude* of the explosive root, not a test for its presence | `rootstamp()` (two S3 methods: default for a single sub-sample, `radf_obj` to run every `datestamp()` episode at once) |
| stands alone | A point-estimation tool, not a test | `contagion_reg()` |

Naming prefixes are a convention, not a contract -- they're easy to
misremember and, as `monitor()` shows, sometimes trade off against each
other (grouped with its fellow monitors under a name that deliberately
doesn't advertise its ADF-family internals, so it can't be mistaken for a
`radf_*()` variant). For anything programmatic, don't parse function
names: call `exuber_functions()`, which returns the same categorization
as actual, queryable data.

```{r registry}
exuber_functions(family = "monitor")
```

Two names that look related but aren't: the `dating_*()` family above are
standalone SSR/BIC procedures called directly on raw data -- they take no
critical value at all. `datestamp()` (see below) is a different thing
entirely: the generic that applies PSY's own threshold-crossing rule to any
`radf_obj` + `radf_cv` pair.

One exception to "`datestamp()` always needs a `radf_cv`": `datestamp(object,
option = "svadf")` runs Sarkar & Wells (2026)'s asymmetric-threshold dating
directly off `object$badf`, no critical value at all -- see
`vignette("experimental-methods")`.

## What actually plugs into `summary()`/`datestamp()`/`tidy()`/`autoplot()`

These four generics are built around one shape: a `radf_obj` (from `radf()`)
paired with a `radf_cv` that carries a **time-varying** boundary
(`badf_cv`/`bsadf_cv`, one critical value per recursion point) as well as the
three scalar sup-statistic critical values (`adf_cv`/`sadf_cv`/`gsadf_cv`).
Only functions whose result actually carries the `radf_obj` class -- and
whose paired `_cv()` actually computes that time-varying boundary -- get the
full pipeline. Three tiers, in practice:

### Full support: `radf_common()`, `radf_kp()`, `radf_tt()`, `radf_sign()`, `radf_sign_dm()`, `radf_sbz()`

`radf_kp()`/`radf_common()` literally return `radf()`'s own output (purged
of volatility, or computed on a PCA factor, respectively, then `radf()`
unmodified), so every generic works exactly as it does for plain `radf()`:

```{r kp-full}
res <- radf_kp(sim_data, minw = 20)
cv <- radf_mc_cv(n = attr(res, "n"), minw = 20)

summary(res, cv = cv)
datestamp(res, cv = cv)
tidy(res, cv = cv)
autoplot(res, cv = cv)
```

The other three are different: they carry the `radf_obj` class but build
their statistic on `gls_dfstat_grid()` (a no-intercept, GLS-demeaned
recursive-DF grid, fed the raw series for `radf_tt()`, its cumulated sign
for `radf_sign()`, a recursively demeaned cumulated sign for
`radf_sign_dm()`) rather than calling `radf()` directly. Until 2026-08-18
all three had a real gap: their `_cv()` functions only ever computed the
three scalar critical values `summary()`/`tidy()` need, discarding the
`badf`/`bsadf` path `gls_dfstat_grid()` already computes per replicate,
so `datestamp()`/`autoplot()` (which need a *time-varying* boundary) always
errored. Fixed for all three the same way, once the pattern was confirmed
in `radf_tt_cv()` first and then checked to hold for the other two as
well: unlike `radf_mc_cv()`'s own `bsadf_cv` (a `cummax()`-across-replicates
shortcut around the base C++ engine's output shape), `gls_dfstat_grid()`'s
`bsadf` is already the genuine sup-over-all-window-starts statistic at
each point, so no shortcut derivation was needed -- just the per-time-point
quantile across replicates, the construction `radf_mc_cv()` uses for its
own `bsadf_cv`. Validated per function: `badf_cv`'s last row is
bit-identical to `adf_cv` (`adf` is literally `badf`'s last point, per
replicate -- a hard identity, not an approximate check, and true regardless
of which series feeds `gls_dfstat_grid()`); empirical false-alarm rate
under `H0` is at or below nominal (`radf_tt` 3.3%, `radf_sign` 5.5%,
`radf_sign_dm` 3.5%, all at nominal 5%, n=100, minw=20); and detection
power on an identical synthetic bubble is in the same range as the
established `radf()`/`radf_mc_cv()` baseline (16%) rather than
suspiciously higher or lower (`radf_tt` 18%, `radf_sign` 20%,
`radf_sign_dm` 8% -- the sign-based tests trading power for their
heteroskedasticity invariance is itself the paper's own documented
finding, not a validation red flag).

```{r tt-full}
res <- radf_tt(sim_data, minw = 20)
cv <- radf_tt_cv(n = 100, minw = 20)

summary(res, cv = cv)
datestamp(res, cv = cv)
tidy(res, cv = cv)
autoplot(res, cv = cv)
```

```{r sign-full}
res <- radf_sign(sim_data, minw = 20)
cv <- radf_sign_cv(n = 100, minw = 20)

summary(res, cv = cv)
datestamp(res, cv = cv)
tidy(res, cv = cv)
autoplot(res, cv = cv)
```

`radf_sbz()` is a fourth, separate case: it builds its statistic (`supBZ`) on
`wls_dfstat_grid()`, a WLS/kernel-volatility-weighted no-intercept
recursive-DF grid, not `gls_dfstat_grid()` -- but the same fix applies for
the same reason, since `wls_dfstat_grid()` already returns the full
`badf`/`bsadf` path per replicate. `radf_sbz_cv()`'s wild bootstrap is
therefore built the same way as `radf_tt_cv()`/`radf_sign_cv()`'s Monte
Carlo simulation: per-time-point quantile across replicates, no `cummax()`
shortcut needed. Validated the same way: `badf_cv`'s last row is
bit-identical to `adf_cv`; empirical false-alarm rate under `H0` is 5.0%
at nominal 5% (n=100, minw=20, 200 replications); and it does reject on a
sufficiently strong deterministic explosive path, though its
kernel-volatility weighting trades away enough power on `sim_data`'s milder
bubbles that none of the five reject at nboot=100-200 -- the same
power/robustness trade-off already documented for `radf_sbz_union()`'s
`supBZ` leg below, not a new finding specific to the split.

```{r sbz-full}
res <- radf_sbz(sim_data, minw = 20)
cv <- radf_sbz_cv(sim_data, minw = 20, nboot = 200, seed = 1)

summary(res, cv = cv)
tidy(res, cv = cv)
```

`datestamp()`/`autoplot()` need at least one rejection to have anything to
show (they error otherwise, same as for any other `radf_obj`/`radf_cv`
pair) -- none of `sim_data`'s five series clear `supBZ`'s bar above, so
here's a series built to:

```{r sbz-full-reject}
set.seed(7)
n <- 120; te <- 70
y <- cumsum(rnorm(n))
y[(te + 1):n] <- y[te] * 1.15 ^ seq_len(n - te)

res2 <- radf_sbz(y, minw = 20)
cv2 <- radf_sbz_cv(y, minw = 20, nboot = 100, seed = 1)
datestamp(res2, cv = cv2)
autoplot(res2, cv = cv2)
```

### No support: everything else

The remaining ~15 functions (`lbi_test()`, `ssu_test()`, `quantile_test()`,
`cobubble_test()`, the `dating_*()` family, the `monitor()`/`monitor_*()`
family (including `monitor()` itself, ADF-family internals notwithstanding --
see above), `contagion_reg()`, `radf_recovery()`, `rootstamp()`,
`radf_sbz_union()`) each return their own class with their own `print()` method,
because their output genuinely doesn't fit the `radf_obj` shape -- a dating
table isn't a per-series sup-statistic, a monitoring alarm isn't a critical
value grid. Trying to force them through `summary()`/`datestamp()`/`tidy()`/
`autoplot()` isn't a documentation gap to close; the right call is their own
presentation, shown directly:

`rootstamp()` is the one exception worth flagging: it's grouped under
**Analysis** in the [reference index](../reference/index.html) and the
README's workflow list, right after `datestamp()`, since that's genuinely
where it belongs in the *sequence of steps* (detect → date → measure
growth rate) -- but that's a workflow position, not an S3-support tier.
It's still its own class with its own `print()` method, same as everything
else in this section; see `vignette("root-inference")`.

```{r standalone}
dating_hls(sim_data$psy1, trim = 0.05)

ssu_test(sim_data$psy1, level = 0.95)
```

## Summary

| Tier | Functions | `summary()` | `datestamp()` | `tidy()` | `autoplot()` |
|---|---|---|---|---|---|
| Full | `radf()`, `radf_common()`, `radf_kp()`, `radf_tt()`, `radf_sign()`, `radf_sign_dm()` | yes | yes | yes | yes |
| Standalone | everything else | own `print()` | -- | -- | -- |
