---
title: "Naming Conventions and the Analysis/Tidying/Plotting Pipeline"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Naming Conventions and the Analysis/Tidying/Plotting Pipeline}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, warning = FALSE, message = FALSE,
  comment = "#>"
)
```

```{r load}
library(exuber)
```

## Why this exists

`exuber` started as one test (`radf()`, the recursive ADF/SADF/GSADF/BSADF
statistic of Phillips, Shi & Yu 2015) and grew, through a long research
programme, into roughly 25 functions covering a dozen papers' worth of
related-but-distinct methodology: alternative tests, dating procedures,
monitoring schemes, root inference. Every one of them used to be named
`radf_<something>()`, which was accurate for some and misleading for others --
a `radf_` prefix reads as "this is a recursive-ADF statistic," but several of
these functions are not that at all. This vignette documents the naming
scheme that replaced it, and -- more usefully -- which functions can actually
be plugged into the shared `summary()`/`datestamp()`/`tidy()`/`autoplot()`
pipeline built for `radf()`, and which have their own, differently-shaped
output instead.

## The naming scheme

| Pattern | Means | Examples |
|---|---|---|
| `radf_` prefix | Genuinely built on the recursive-ADF core: calls `radf()` directly, or reuses its `badf`/`bsadf` recursion | `radf()`, `radf_tt()`, `radf_sign()`, `radf_common()`, `radf_kp()`, `radf_recovery()`, `radf_svadf()`, and the `_cv`/`_mc`/`_sb`/`_wb` critical-value engines |
| `_test` suffix | A standalone hypothesis test with its own null distribution, not built on the recursive-ADF core | `lbi_test()`, `ssu_test()`, `quantile_test()`, `cobubble_test()` |
| `dating_` prefix | Point-estimation / model-selection dating, no formal hypothesis test at all | `dating_hls()`, `dating_hlw()`, `dating_knp()`, `dating_pdc()` |
| `monitor_` prefix | Real-time/sequential detection -- grouped by *what it does*, not by internal mechanism, so this one deliberately crosses the ADF-family line: `monitor_radf()` reuses `badf`/`bsadf` directly (same as the `radf_` family above) but is named for its monitoring behavior, not that internal detail, since that's what a user searching for "what can monitor a series in real time" actually wants to find | `monitor_cusum()`, `monitor_lbi()`, `monitor_quantile()`, `monitor_radf()` |
| `root_` prefix | Confidence-interval inference on the *magnitude* of the explosive root, not a test for its presence | `rootstamp()` (two S3 methods: default for a single sub-sample, `radf_obj` to run every `datestamp()` episode at once) |
| stands alone | A point-estimation tool, not a test | `contagion_reg()` |

Naming prefixes are a convention, not a contract -- they're easy to
misremember and, as `monitor_radf()` shows, sometimes trade off against each
other (grouped with its fellow monitors here, at the cost of no longer
signaling its ADF-family internals). For anything programmatic, don't parse
function names: call `exuber_functions()`, which returns the same
categorization as actual, queryable data.

```{r registry}
exuber_functions(family = "monitor")
```

Two names that look related but aren't: the `dating_*()` family above are
standalone SSR/BIC procedures called directly on raw data -- they take no
critical value at all. `datestamp()` (see below) is a different thing
entirely: the generic that applies PSY's own threshold-crossing rule to any
`radf_obj` + `radf_cv` pair.

## What actually plugs into `summary()`/`datestamp()`/`tidy()`/`autoplot()`

These four generics are built around one shape: a `radf_obj` (from `radf()`)
paired with a `radf_cv` that carries a **time-varying** boundary
(`badf_cv`/`bsadf_cv`, one critical value per recursion point) as well as the
three scalar sup-statistic critical values (`adf_cv`/`sadf_cv`/`gsadf_cv`).
Only functions whose result actually carries the `radf_obj` class -- and
whose paired `_cv()` actually computes that time-varying boundary -- get the
full pipeline. Three tiers, in practice:

### Full support: `radf_common()`, `radf_kp()`

Both literally return `radf()`'s own output (`radf_kp()` purges volatility
first, `radf_common()` extracts a PCA factor first, then calls `radf()`
unmodified), so every generic works exactly as it does for plain `radf()`:

```{r kp-full}
res <- radf_kp(sim_data, minw = 20)
cv <- radf_mc_cv(n = attr(res, "n"), minw = 20)

summary(res, cv = cv)
datestamp(res, cv = cv)
tidy(res, cv = cv)
autoplot(res, cv = cv)
```

### Partial support: `radf_sign()`, `radf_sign_dm()`, `radf_tt()`

These carry the `radf_obj` class too (so `summary()` and `tidy()` work), but
their own critical-value functions -- `radf_sign_cv()`, `radf_sign_dm_cv()`,
`radf_tt_cv()` -- only ever computed the three scalar critical values
`summary()` needs, never the `badf_cv`/`bsadf_cv` time-varying boundary
`datestamp()`/`autoplot()` require. This is a real, currently unfixed gap,
not a design choice -- adding it means simulating and calibrating a full
recursive boundary the way `radf_mc_cv()`/`radf_common_cv()` do, which needs
this package's usual validation pass (formula-exact check, Monte Carlo size,
a replication script) before it ships, not a quick patch.

```{r sign-partial}
res <- radf_sign(sim_data, minw = 20)
cv <- radf_sign_cv(n = 100, minw = 20)

summary(res, cv = cv) # works
tidy(res, cv = cv) # works
```
```r
datestamp(res, cv = cv) # errors: cv has no bsadf_cv
autoplot(res, cv = cv) # errors: cv has no bsadf_cv
```

### No support: everything else

The remaining ~15 functions (`lbi_test()`, `ssu_test()`, `quantile_test()`,
`cobubble_test()`, the `dating_*()` family, the `monitor_*()` family
(including `monitor_radf()`, ADF-family internals notwithstanding --
see above), `contagion_reg()`, `radf_recovery()`, `radf_svadf()`,
`radf_sbz_cv()`) each return their own class with their own `print()` method,
because their output genuinely doesn't fit the `radf_obj` shape -- a dating
table isn't a per-series sup-statistic, a monitoring alarm isn't a critical
value grid. Trying to force them through `summary()`/`datestamp()`/`tidy()`/
`autoplot()` isn't a documentation gap to close; the right call is their own
presentation, shown directly:

```{r standalone}
dating_hls(sim_data$psy1, trim = 0.05)

ssu_test(sim_data$psy1, level = 0.95)
```

## Summary

| Tier | Functions | `summary()` | `datestamp()` | `tidy()` | `autoplot()` |
|---|---|---|---|---|---|
| Full | `radf()`, `radf_common()`, `radf_kp()` | yes | yes | yes | yes |
| Partial | `radf_sign()`, `radf_sign_dm()`, `radf_tt()` | yes | no (gap) | yes | no (gap) |
| Standalone | everything else | own `print()` | -- | -- | -- |
