---
title: "Naming Conventions and the Analysis/Tidying/Plotting Pipeline"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Naming Conventions and the Analysis/Tidying/Plotting Pipeline}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, warning = FALSE, message = FALSE,
  comment = "#>"
)
```

```{r load}
library(exuber)
```

## Why this exists

`exuber` started as a single test, `radf()`, the recursive ADF/SADF/GSADF/BSADF statistic of Phillips, Shi & Yu (2015). Through a long research programme it grew to roughly 25 functions that cover a dozen papers: alternative tests, dating procedures, monitoring schemes and root inference. Every one of them used to be named `radf_<something>()`. That was accurate for some and misleading for others, because a `radf_` prefix suggests a recursive ADF statistic and several of these functions are not one. This vignette documents the naming scheme that replaced the old one. It also explains which functions plug into the `summary()`, `datestamp()`, `tidy()` and `autoplot()` pipeline built for `radf()`, and which have differently shaped output of their own.

## The naming scheme

| Pattern | Means | Examples |
|---|---|---|
| `radf_` prefix | Built on the recursive ADF core: calls `radf()` directly or reuses its `badf`/`bsadf` recursion | `radf()`, `radf_tt()`, `radf_sign()`, `radf_common()`, `radf_kp()`, `radf_recovery()`, and the `_cv`/`_mc`/`_sb`/`_wb` critical-value engines |
| `_test` suffix | A standalone hypothesis test with its own null distribution, not built on the recursive ADF core | `lbi_test()`, `ssu_test()`, `quantile_test()`, `cobubble_test()` |
| `dating_` prefix | Dating by point estimation and model selection, with no formal hypothesis test | `dating_hls()`, `dating_hlw()`, `dating_knp()`, `dating_pdc()` |
| `monitor`/`monitor_` prefix | Real-time (sequential) detection, grouped by what the function does and not by its internal mechanism. `monitor()` is the flagship of this family, as `radf()` is for the `radf_` family. It reuses `badf` and `bsadf` directly but carries no `radf` or `sadf` token, so that it reads as the real-time monitor and not as a `radf_` variant (see below) | `monitor()`, `monitor_cusum()`, `monitor_lbi()`, `monitor_quantile()` |
| `root` family | Confidence-interval inference on the magnitude of the explosive root, not a test for its presence (`exuber_functions(family = "root")`; there is no shared prefix) | `rootstamp()`, which has two S3 methods: the default for a single sub-sample and the `radf_obj` method, which runs every `datestamp()` episode at once |
| stands alone | A point-estimation tool, not a test | `contagion_reg()` |

The prefixes are a convention and not a contract. They are easy to misremember and they sometimes pull against each other. `monitor()` is grouped with the other monitors under a name that deliberately does not advertise its ADF-family internals, so that nobody mistakes it for a `radf_*()` variant. For programmatic use, do not parse function names. Call `exuber_functions()`, which returns the same categorization as queryable data.

```{r registry}
exuber_functions(family = "monitor")
```

Two names look related but are not. The `dating_*()` functions above are standalone SSR/BIC procedures that run directly on raw data and take no critical value. `datestamp()` (see below) is a different thing: it is the generic that applies the threshold-crossing rule of Phillips, Shi and Yu to any `radf_obj` and `radf_cv` pair.

`datestamp()` normally needs a `radf_cv`, with one exception. `datestamp(object, option = "svadf")` runs the asymmetric-threshold dating of Sarkar & Wells (2026) directly on `object$badf` and needs no critical value (see `vignette("experimental-methods")`).

## What actually plugs into `summary()`/`datestamp()`/`tidy()`/`autoplot()`

These four generics are built around one shape: a `radf_obj` (from `radf()`) paired with a `radf_cv` that carries a time-varying boundary (`badf_cv` and `bsadf_cv`, one critical value per recursion point) as well as the three scalar sup-statistic critical values (`adf_cv`, `sadf_cv` and `gsadf_cv`). Only functions whose result has the `radf_obj` class, and whose paired `_cv()` function computes that time-varying boundary, get the full pipeline. In practice there are three tiers.

### Full support: `radf_common()`, `radf_kp()`, `radf_tt()`, `radf_sign()`, `radf_sign_dm()`, `radf_sbz()`

`radf_kp()` and `radf_common()` return the output of `radf()` itself, computed on a series purged of volatility or on a PCA factor respectively, so every generic works exactly as it does for plain `radf()`. The running example for this section is the series these tests were designed for: the `sim_psy1()` bubble with a permanent volatility break (`sim_vol_break()`, where the innovation standard deviation triples half-way through the sample). See `vignette("volatility-robust-radf")`.

```{r vol-dgp}
y <- sim_psy1(n = 200, seed = 1, e = sim_vol_break(199))
```

```{r kp-full}
res <- radf_kp(y, minw = 20)
cv <- radf_mc_cv(n = attr(res, "n"), minw = 20)

summary(res, cv = cv)
datestamp(res, cv = cv)
tidy(res, cv = cv)
autoplot(res, cv = cv)
```

The other three are different. They carry the `radf_obj` class but build their statistic on `gls_dfstat_grid()` and do not call `radf()` directly. This function is a no-intercept, GLS-demeaned recursive Dickey-Fuller grid. It is fed the raw series for `radf_tt()`, the cumulated sign of the series for `radf_sign()`, and a recursively demeaned cumulated sign for `radf_sign_dm()`.

Until 2026-08-18 the `_cv()` functions of all three had a gap. They computed only the three scalar critical values that `summary()` and `tidy()` need and discarded the `badf` and `bsadf` paths that `gls_dfstat_grid()` already produces for each replicate. As a result `datestamp()` and `autoplot()`, which need a time-varying boundary, always failed. We fixed all three in the same way. We first established the fix in `radf_tt_cv()` and then checked that it holds for the other two. The `bsadf` that `gls_dfstat_grid()` returns is already the sup over all window starts at each point. This differs from the `bsadf_cv` of `radf_mc_cv()`, which uses a `cummax()` across replicates because of the output shape of the base C++ engine. No shortcut was needed here, and the boundary is the per-time-point quantile across replicates, the same construction `radf_mc_cv()` uses for its own `bsadf_cv`.

We validated each function in three ways.

- The last row of `badf_cv` is bit-identical to `adf_cv`, because `adf` is the last point of `badf` in every replicate. This is an exact identity, and it holds whichever series feeds `gls_dfstat_grid()`.
- The empirical false-alarm rate under `H0` is at or below the nominal 5% (`radf_tt` 3.3%, `radf_sign` 5.5%, `radf_sign_dm` 3.5%, with n = 100 and minw = 20).
- The detection power on an identical synthetic bubble is in the same range as the 16% of the established `radf()` and `radf_mc_cv()` baseline, and it is neither suspiciously higher nor lower (`radf_tt` 18%, `radf_sign` 20%, `radf_sign_dm` 8%). The sign-based tests give up power in exchange for invariance to heteroskedasticity, which is a documented finding of the source paper and not a validation problem.

```{r tt-full}
res <- radf_tt(y, minw = 20)
cv <- radf_tt_cv(n = 200, minw = 20)

summary(res, cv = cv)
datestamp(res, cv = cv)
tidy(res, cv = cv)
autoplot(res, cv = cv)
```

```{r sign-full}
res <- radf_sign(y, minw = 20)
cv <- radf_sign_cv(n = 200, minw = 20)

summary(res, cv = cv)
datestamp(res, cv = cv)
tidy(res, cv = cv)
autoplot(res, cv = cv)
```

`radf_sbz()` is a fourth, separate case. It builds its statistic (`supBZ`) on `wls_dfstat_grid()`, a no-intercept recursive Dickey-Fuller grid weighted by WLS and kernel volatility, and not on `gls_dfstat_grid()`. The same fix applies for the same reason, since `wls_dfstat_grid()` already returns the full `badf` and `bsadf` path for each replicate. The wild bootstrap in `radf_sbz_cv()` is therefore built as the Monte Carlo simulation in `radf_tt_cv()` and `radf_sign_cv()` is, with a per-time-point quantile across replicates and no `cummax()` shortcut. We validated it in the same way. The last row of `badf_cv` is bit-identical to `adf_cv`, and the empirical false-alarm rate under `H0` is 5.0% at a nominal 5% (n = 100, minw = 20, 200 replications). The test also rejects on a sufficiently strong deterministic explosive path. Its kernel-volatility weighting costs enough power, however, that it does not reject the series above at nboot = 100 to 200, where the bubble is the milder default of `sim_psy1()`. The same trade between power and robustness is documented for the `supBZ` leg of `radf_sbz_union()` below, so it is not new to this split.

```{r sbz-full}
res <- radf_sbz(y, minw = 20)
cv <- radf_sbz_cv(y, minw = 20, nboot = 200, seed = 1)

summary(res, cv = cv)
tidy(res, cv = cv)
```

`datestamp()` and `autoplot()` need at least one rejection to have anything to show, and they raise an error otherwise, as for any other `radf_obj` and `radf_cv` pair. The series above does not clear the `supBZ` threshold, so we repeat the volatility break with a stronger explosive regime that does not collapse (`rho = 1.03` from `t = 120` to the end of the sample):

```{r sbz-full-reject}
y_strong <- sim_psy1(n = 200, te = 120, tf = 200, c = 0.03, alpha = 0, seed = 1,
                     e = sim_vol_break(199))

res2 <- radf_sbz(y_strong, minw = 20)
cv2 <- radf_sbz_cv(y_strong, minw = 20, nboot = 100, seed = 1)
datestamp(res2, cv = cv2)
autoplot(res2, cv = cv2)
```

### Own `print()` and `autoplot()`: everything else

The remaining 15 or so functions return their own class, with their own `print()` and `autoplot()` methods. They are `lbi_test()`, `ssu_test()`, `quantile_test()`, `cobubble_test()`, the `dating_*()` family, the `monitor()` and `monitor_*()` family (including `monitor()` itself, despite its ADF-family internals), `contagion_reg()`, `radf_recovery()`, `rootstamp()` and `radf_sbz_union()`. Their output does not fit the `radf_obj` shape: a dating table is not a per-series sup-statistic, and a monitoring alarm is not a critical value grid. Forcing them through `summary()`, `datestamp()` and `tidy()` would not close a documentation gap, so each is presented by its own methods, shown below.

`rootstamp()` needs one remark. The [reference index](../reference/index.html) and the workflow list in the README place it under Analysis, right after `datestamp()`, because that is its position in the sequence of steps (detect, date, measure the growth rate). That is a position in the workflow and not an S3-support tier. It has its own class with its own `print()` and `autoplot()` methods, like everything else in this section. See `vignette("root-inference")`.

```{r standalone}
dating_hls(sim_data$psy1, trim = 0.05)

ssu_test(sim_data$psy1, sig_lvl = 95)
autoplot(dating_hls(sim_data$psy1, trim = 0.05))
```

## Summary

| Tier | Functions | `summary()` | `datestamp()` | `tidy()` | `autoplot()` |
|---|---|---|---|---|---|
| Full | `radf()`, `radf_common()`, `radf_kp()`, `radf_tt()`, `radf_sign()`, `radf_sign_dm()`, `radf_sbz()` | yes | yes | yes | yes |
| Standalone | everything else | own `print()` | -- | -- | own method |
