---
title: "Plotting with exuber"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plotting with exuber}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, warning = FALSE, message = FALSE,
  comment = "#>", fig.width = 8, fig.height = 4.5
)
```

```{r load}
library(exuber)
library(ggplot2)
```

## One `autoplot()` per object

Every result object in the package has an `autoplot()` method. The call is always `autoplot(x)`, and what it draws depends on the class of `x`. Each method returns an ordinary `ggplot` object, so anything ggplot2 offers, such as themes, scales, extra layers and `facet_*()` arguments, can be added on top.

| What you have | What `autoplot()` draws |
|---|---|
| `radf()` result (`radf_obj`) | Test statistic against the critical-value sequence, with one facet per series that rejects the null and the explosive episodes shaded |
| `datestamp()` result (`ds_radf`) | Only the episodes, as one horizontal segment per series. This is the view for many series at once |
| `radf_*_distr()` result (`radf_distr`) | The simulated null distribution of the ADF/SADF/GSADF statistics |
| Any `sim_*()` series | The series itself |
| `monitor()`, `monitor_*()` | Monitored statistic against its boundary, with markers for the end of training and for the alarm |
| `dating_*()`, `radf_recovery()` | The series with vertical markers at the estimated break dates |
| `rootstamp()` | Estimated root and its confidence interval per episode |
| `lbi_test()`, `quantile_test()`, `radf_sbz_union()` | Statistic against critical value for each series |

The rest of this vignette covers the first three rows, which belong to the `radf()` workflow, and shows how to build your own plot from the tidied tables when the defaults do not fit. The other methods take no options beyond the object. They appear in their own vignettes (`vignette("monitoring")`, `vignette("dating-methods")` and `vignette("root-inference")`).

## The `radf()` plot

We simulate four series, one from each of the classic bubble data generating processes in the package (see `vignette("simulation")`), and estimate them with one lag. The critical values depend on `(n, lag)`, so we simulate them once and pass them to every call, as in `vignette("exuber")`:

```{r data}
sims <- data.frame(
  psy1 = sim_psy1(100, seed = 1),
  psy2 = sim_psy2(100, seed = 2),
  evans = sim_evans(100, seed = 3),
  blan = sim_blan(100, seed = 4)
)
est <- radf(sims, lag = 1)
cv <- radf_mc_cv(100, lag = 1, seed = 1)
```

```{r autoplot-basic}
autoplot(est, cv)
```

Only series that reject the null at the 5% level are drawn. The arguments of `autoplot()` itself control what is plotted:

```{r autoplot-options, fig.height = 3}
# Every series, whether or not it rejects
autoplot(est, cv, nonrejected = TRUE)

# A subset, by name or position; the SADF sequence instead of the BSADF one
autoplot(est, cv, select_series = c("psy1", "evans"), option = "sadf")
```

The shading of the explosive episodes is a `geom_rect()` layer. The `shade_opt` argument and the `shade()` helper control it, and `shade_opt = NULL` removes it:

```{r autoplot-shade, fig.height = 3}
autoplot(est, cv, select_series = "psy2",
         shade_opt = shade(fill = "pink", opacity = 0.3))
```

`autoplot2()` draws the series itself instead of the statistic, with the same shading. This is often easier to read for a non-technical audience. The method for `datestamp()` objects reduces each series to its episodes:

```{r autoplot2, fig.height = 3}
autoplot2(est, cv, select_series = "psy2")
```

```{r autoplot-ds, fig.height = 2}
datestamp(est, cv) %>%
  autoplot()
```

### Changing the appearance

Colors, line types and themes are handled by ggplot2. `autoplot()` maps the statistic and the critical value to `color`, `size` and `linetype`, so the ggplot2 `scale_*_manual()` functions can override them. `scale_exuber_manual()` sets all three at once. `theme_exuber()` is the default theme of the package, and it is exported so that you can apply it to your own plots too:

```{r autoplot-theme, fig.height = 3}
autoplot(est, cv, select_series = "psy2") +
  scale_exuber_manual(color_values = c("grey40", "black"),
                      linetype_values = c(3, 1)) +
  theme_classic()
```

Arguments that `autoplot()` does not recognize are passed on to `ggplot2::facet_wrap()`, so `scales = "free_y"`, `ncol` and `labeller` work directly. `?autoplot.radf_obj` has a labeller example that renames the facets.

## Building your own plot

When the default layout does not suit you, skip `autoplot()` and start from the table it is built on. `augment_join()` joins the full statistic sequences of a `radf_obj` with the critical-value sequences of a `radf_cv`. It returns one row per observation, series, statistic and significance level, which ggplot2 can use as it is:

```{r augment-join}
joined <- augment_join(est, cv)
joined
```

```{r custom-facet, fig.height = 6}
joined %>%
  ggplot(aes(x = index)) +
  geom_line(aes(y = tstat)) +
  geom_line(aes(y = crit), linetype = 2) +
  facet_grid(sig + stat ~ id, scales = "free_y") +
  theme_exuber()
```

`tidy_join()` is the scalar counterpart, with one row per series and statistic, which is the table that `summary()` prints. Calling `tidy()` or `augment()` on either object alone returns the two halves before they are joined. `vignette("naming-and-analysis")` describes the full pipeline.

## Distributions

The `radf_*_distr()` functions are the counterparts of the critical-value functions. They return the whole simulated null distribution instead of its quantiles, and they have their own `autoplot()` method:

```{r distr}
distr <- radf_mc_distr(n = 100, nrep = 1000, seed = 1)
autoplot(distr)
```

As elsewhere, `tidy()` returns the underlying table, so an empirical CDF or any other summary takes only a few lines of ggplot2:

```{r ecdf}
distr %>%
  tidy() %>%
  tidyr::pivot_longer(everything(), names_to = "statistic") %>%
  ggplot(aes(value, color = statistic)) +
  stat_ecdf() +
  geom_hline(yintercept = 0.95, linetype = 2) +
  labs(title = "Empirical CDF of the null distributions", y = NULL) +
  theme_exuber()
```

## Which to reach for

- For a quick look at which series are explosive and when, use `autoplot(est, cv)`. Add `nonrejected = TRUE` to include the series that do not reject.
- To show the series itself with the episodes shaded, for a non-technical reader, use `autoplot2(est, cv)`.
- To show only the episodes of many series, use `autoplot(datestamp(est, cv))`.
- For cosmetic changes, keep `autoplot()` and add ggplot2 layers, scales or a theme. Use `scale_exuber_manual()` to style the statistic and the critical value.
- For a different layout altogether, build your own plot from `augment_join(est, cv)`.
