---
title: "Plotting with exuber"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plotting with exuber}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, warning = FALSE, message = FALSE,
  comment = "#>", fig.width = 8, fig.height = 4.5
)
```

```{r load}
library(exuber)
library(ggplot2)
```

## One `autoplot()` per object

Every result object in the package has an `autoplot()` method, so the
plotting call is the same whatever you ran -- `autoplot(x)` -- and what it
draws depends on what `x` is. All of them return a plain `ggplot` object,
so anything ggplot2 can do (themes, scales, extra layers, `facet_*()`
arguments) applies on top.

| What you have | What `autoplot()` draws |
|---|---|
| `radf()` result (`radf_obj`) | Test statistic vs. critical-value sequence, one facet per series that rejects the null, with the explosive episodes shaded |
| `datestamp()` result (`ds_radf`) | Just the episodes, one horizontal segment per series -- the view for many series at once |
| `radf_*_distr()` result (`radf_distr`) | The simulated null distribution of the ADF/SADF/GSADF statistics |
| Any `sim_*()` series | The series itself |
| `monitor()`, `monitor_*()` | Monitored statistic vs. its boundary, with training-end and alarm markers |
| `dating_*()`, `radf_recovery()` | The series with vertical markers at the estimated break dates |
| `rootstamp()` | Estimated root and its confidence interval per episode |
| `lbi_test()`, `quantile_test()`, `radf_sbz_union()` | Statistic vs. critical value per series |

The rest of this vignette is about the first three -- the `radf()`
workflow -- and about building your own plot from the tidied tables when
the defaults don't fit. The other methods take no options beyond the
object and are shown in their own vignettes (`vignette("monitoring")`,
`vignette("dating-methods")`, `vignette("root-inference")`).

## The `radf()` plot

Four simulated series, one from each of the package's classic bubble
DGPs (see `vignette("simulation")`), estimated with one lag. With a lag
the critical values depend on `(n, lag)`, so we simulate them once and
pass them to every call -- the same pattern as `vignette("exuber")`:

```{r data}
sims <- data.frame(
  psy1 = sim_psy1(100, seed = 1),
  psy2 = sim_psy2(100, seed = 2),
  evans = sim_evans(100, seed = 3),
  blan = sim_blan(100, seed = 4)
)
est <- radf(sims, lag = 1)
cv <- radf_mc_cv(100, lag = 1, seed = 1)
```

```{r autoplot-basic}
autoplot(est, cv)
```

Only series that reject the null at 5% are drawn. The options that change
*what* is plotted are arguments of `autoplot()` itself:

```{r autoplot-options, fig.height = 3}
# Every series, whether or not it rejects
autoplot(est, cv, nonrejected = TRUE)

# A subset, by name or position; the SADF sequence instead of the BSADF one
autoplot(est, cv, select_series = c("psy1", "evans"), option = "sadf")
```

The shading of the explosive episodes is a `geom_rect()` layer controlled
through `shade_opt` and the `shade()` helper; `shade_opt = NULL` removes it:

```{r autoplot-shade, fig.height = 3}
autoplot(est, cv, select_series = "psy2",
         shade_opt = shade(fill = "pink", opacity = 0.3))
```

`autoplot2()` draws the *series* instead of the statistic, with the same
shading, which is often the more readable view for a non-technical
audience; `datestamp()`'s own method reduces each series to its episodes:

```{r autoplot2, fig.height = 3}
autoplot2(est, cv, select_series = "psy2")
```

```{r autoplot-ds, fig.height = 2}
datestamp(est, cv) %>%
  autoplot()
```

### Changing the appearance

Everything else -- colors, line types, theme -- is ggplot2 territory.
`autoplot()` maps the statistic and the critical value to `color`, `size`
and `linetype`, so the ggplot2 `scale_*_manual()` functions override
them; `scale_exuber_manual()` sets all three at once, and `theme_exuber()`
is the package's default theme, exported so you can apply it to your own
plots too:

```{r autoplot-theme, fig.height = 3}
autoplot(est, cv, select_series = "psy2") +
  scale_exuber_manual(color_values = c("grey40", "black"),
                      linetype_values = c(3, 1)) +
  theme_classic()
```

Arguments `autoplot()` doesn't recognize are forwarded to
`ggplot2::facet_wrap()`, so `scales = "free_y"`, `ncol`, `labeller`, etc.
work directly -- see `?autoplot.radf_obj` for a labeller example that
renames the facets.

## Building your own plot

When the default layout isn't what you need, skip `autoplot()` and start
from the same table it is built on. `augment_join()` joins the full
statistic sequences of a `radf_obj` with the critical-value sequences of a
`radf_cv`, one row per observation, series, statistic and significance
level -- ggplot2-ready as is:

```{r augment-join}
joined <- augment_join(est, cv)
joined
```

```{r custom-facet, fig.height = 6}
joined %>%
  ggplot(aes(x = index)) +
  geom_line(aes(y = tstat)) +
  geom_line(aes(y = crit), linetype = 2) +
  facet_grid(sig + stat ~ id, scales = "free_y") +
  theme_exuber()
```

`tidy_join()` is the scalar counterpart (one row per series and statistic,
the table `summary()` prints), and `tidy()`/`augment()` on either object
alone give the un-joined halves -- see `vignette("naming-and-analysis")`
for the full pipeline.

## Distributions

The `radf_*_distr()` twins of the critical-value functions return the
whole simulated null distribution rather than its quantiles, and have
their own `autoplot()`:

```{r distr}
distr <- radf_mc_distr(n = 100, nrep = 1000, seed = 1)
autoplot(distr)
```

As with everything else, `tidy()` gives the underlying table, so an
empirical CDF or any other summary is a few ggplot2 lines away:

```{r ecdf}
distr %>%
  tidy() %>%
  tidyr::pivot_longer(everything(), names_to = "statistic") %>%
  ggplot(aes(value, color = statistic)) +
  stat_ecdf() +
  geom_hline(yintercept = 0.95, linetype = 2) +
  labs(title = "Empirical CDF of the null distributions", y = NULL) +
  theme_exuber()
```

## Which to reach for

- A quick look at which series are explosive and when:
  `autoplot(est, cv)`; add `nonrejected = TRUE` to see the ones that
  aren't.
- The series itself with the episodes shaded, for a non-technical reader:
  `autoplot2(est, cv)`.
- Many series, only the episodes: `autoplot(datestamp(est, cv))`.
- Cosmetic changes: keep `autoplot()` and add ggplot2 layers, scales or a
  theme; `scale_exuber_manual()` for the statistic/critical-value styling.
- A different layout altogether: `augment_join(est, cv)` and build it
  yourself.
