---
title: "Time-Transformed Test (STADF/GSTADF)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Time-Transformed Test (STADF/GSTADF)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, warning = FALSE, message = FALSE,
  comment = "#>"
)
```

```{r load}
library(exuber)
```

## Why another test

`radf()` (the classic PSY GSADF test) assumes the innovation variance is
constant. Real financial series usually don't have constant volatility, and
Harvey, Leybourne, Sollis & Taylor (2016) show that when volatility is
time-varying, `radf()`'s standard critical values no longer control size --
`radf_wb_cv()` already addresses this in exuber via a wild bootstrap.

`radf_tt()` implements a different, bootstrap-free fix from Kurozumi,
Skrobotov & Tsarev (2024, *Journal of Financial Econometrics*): instead of
resampling, it *time-deforms* the series using a nonparametric estimate of
its variance profile, so that the deformed series behaves like a
constant-volatility random walk under the null. The resulting statistic's
null distribution is then the same (pivotal) distribution as under
homoskedasticity, so ordinary asymptotic critical values apply -- no
bootstrap, and no per-dataset resimulation needed.

## Basic usage

```{r radf-tt}
set.seed(1)
y <- sim_psy1(100)
res <- radf_tt(y)
res
```

`radf_tt_cv()` gives the matching (pivotal) asymptotic critical values;
because the null distribution doesn't depend on the volatility path, one
call with a large `n` approximates the whole family, unlike
`radf_wb_cv()`'s per-dataset bootstrap:

```{r radf-tt-cv}
cv <- radf_tt_cv(n = 300, minw = 30, nrep = 1000, seed = 1)
cv$gsadf_cv
```

## What's actually estimated

Under the hood, `radf_tt()`:

1. estimates the time-varying AR(1) coefficient with a local kernel
   regression, and from its (truncated) residuals builds a monotone
   *variance profile* `eta_hat(s)`, `s` in `[0, 1]`;
2. inverts it and uses the inverse to resample/time-deform the series;
3. runs a (GLS-demeaned, no-intercept) recursive sup-ADF statistic on the
   deformed series -- the same statistic family as `radf()`, but built to
   need no fitted intercept, matching the paper's derivation.

`kernel` (`"uniform"`, the paper's own choice, or `"gaussian"`) and `h`
(bandwidth; default a fixed plug-in, not the paper's full cross-validation
search -- see the package's enhancement notes for the cost/benefit
reasoning) can both be adjusted.

## Verifying against the paper

Kurozumi, Skrobotov & Tsarev's footnote 4 gives an exact published
asymptotic critical value triple for `minw/n = 0.1`: `(2.319, 2.626,
3.223)` at the (10%, 5%, 1%) levels -- for the **STADF** statistic (the
single-sup, `r1 = 0` case). exuber's test suite
(`tests/testthat/test-tt.R`) reproduces this via `radf_tt_cv()`'s own
Monte Carlo and checks it lands within Monte Carlo/finite-sample tolerance
of the published numbers.
