---
title: "Root Inference: How Fast Is the Bubble Growing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Root Inference: How Fast Is the Bubble Growing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, warning = FALSE, message = FALSE,
  comment = "#>"
)
```

```{r load}
library(exuber)
```

## A different question from "is there a bubble"

`radf()`/`datestamp()` (and the `dating_*()`/`_test()`/`monitor()`/`monitor_*()`
families) all answer some version of "is there a bubble, and when did it
happen." None of them
say anything about its *magnitude*: once an explosive episode is dated, how
fast is the underlying autoregressive root actually growing? `rootstamp()`
(Phillips & Magdalinos 2007; Guo, Sun & Wang 2019) answers exactly that, as
a follow-up step *after* detection and dating, not a replacement for either.

It fits a no-intercept AR(1), `y_t = rho * y_{t-1} + e_t`, over a given
sub-sample and reports `rho`'s estimate together with a confidence interval
and an implied **doubling time** (`log(2) / log(rho)`: how many periods
until the bubble doubles in size at the estimated growth rate). Two
methods, for two different starting points -- a plain numeric sub-sample
(fit once, one CI), or a `radf_obj` plus its `datestamp()` result (fit
every episode at once, no manual loop). Neither method's return value
carries the `radf_obj` class itself, so `rootstamp()` doesn't plug into
`summary()`/`tidy()`/`autoplot()` -- see `vignette("naming-and-analysis")`.

## Detect, date, then estimate the root

A unit-root run-up followed by a genuine explosive regime with `rho = 1.04`:

```{r dgp}
set.seed(2026)
burn <- cumsum(rnorm(60))
bubble <- burn[length(burn)] * 1.04^(1:40) + cumsum(rnorm(40, sd = 0.5))
y <- c(burn, bubble)
```

Detect and date it first, the ordinary way:

```{r detect}
r <- radf(y, minw = 20)
cv <- radf_mc_cv(length(y), minw = 20, nrep = 300, seed = 4)
ds <- datestamp(r, cv = cv, min_duration = 3)
ds
```

Then estimate the root over the detected episode -- the default method
takes the sub-sample directly, sliced by the episode's own `Start`/`End`:

```{r root}
ep <- ds[["series1"]]
rootstamp(y[ep$Start[1]:ep$End[1]]) # normal-t interval (Guo, Sun & Wang 2019), true rho = 1.04
rootstamp(y[ep$Start[1]:ep$End[1]], type = "cauchy") # fixed-root Cauchy interval (Phillips & Magdalinos 2007)
```

`rho` recovers something close to the true 1.04, alongside `rho_ci` and the
implied `doubling_time`/`doubling_time_ci`. The two interval types answer
slightly different questions. `type = "normal"` (the default) is the safer
choice under drift/weak dependence and gives a noticeably tighter interval
here; `type = "cauchy"` assumes a genuinely *fixed*, non-drifting root and,
because a Cauchy distribution has much fatter tails than a normal, produces
a visibly wider interval even at the same nominal level.

## Every episode at once

For a `datestamp()` result with more than one episode, the `radf_obj`
method runs the default method on each one without a manual loop -- pass
the original `radf()` result and the `datestamp()` result together:

```{r rootstamp-radf-obj}
rootstamp(r, ds)
```

Root inference on a very short episode is close to meaningless (too few
points to estimate an AR(1) coefficient precisely) -- filter with
`datestamp(..., min_duration = ...)` before piping in, rather than
expecting this method to second-guess what counts as "too short."
