---
title: "Root Inference: How Fast Is the Bubble Growing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Root Inference: How Fast Is the Bubble Growing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, warning = FALSE, message = FALSE,
  comment = "#>"
)
```

```{r load}
library(exuber)
```

## A different question from "is there a bubble"

`radf()`/`datestamp()` (and the `dating_*()`/`_test()` families) all answer
some version of "is there a bubble, and when did it happen." None of them
say anything about its *magnitude*: once an explosive episode is dated, how
fast is the underlying autoregressive root actually growing? `explosive_root()`,
`root_ci()` and `root_ci_datestamp()` (Phillips & Magdalinos 2007; Guo, Sun &
Wang 2019) answer exactly that, as a follow-up step *after* detection and
dating, not a replacement for either.

They fit a no-intercept AR(1), `y_t = rho * y_{t-1} + e_t`, over a given
sub-sample and report `rho`'s estimate together with a confidence interval
and an implied **doubling time** (`log(2) / log(rho)`: how many periods
until the bubble doubles in size at the estimated growth rate). None of the
three carry the `radf_obj` class -- see `vignette("naming-and-analysis")`.

## Detect, date, then estimate the root

A unit-root run-up followed by a genuine explosive regime with `rho = 1.04`:

```{r dgp}
set.seed(2026)
burn <- cumsum(rnorm(60))
bubble <- burn[length(burn)] * 1.04^(1:40) + cumsum(rnorm(40, sd = 0.5))
y <- c(burn, bubble)
```

Detect and date it first, the ordinary way:

```{r detect}
r <- radf(y, minw = 20)
cv <- radf_mc_cv(length(y), minw = 20, nrep = 300, seed = 4)
ds <- datestamp(r, cv = cv, min_duration = 3)
ds
```

Then estimate the root over the detected episode:

```{r root}
est <- explosive_root(y, ds[["series1"]]$Start[1], ds[["series1"]]$End[1])
est
```

`est$rho` recovers something close to the true 1.04. `root_ci()` turns the
point estimate into an interval and a doubling time:

```{r ci}
root_ci(est) # normal-t interval (Guo, Sun & Wang 2019)
root_ci(est, type = "cauchy") # fixed-root Cauchy interval (Phillips & Magdalinos 2007)
```

The two interval types answer slightly different questions. `type = "normal"`
(the default) is the safer choice under drift/weak dependence and gives a
noticeably tighter interval here; `type = "cauchy"` assumes a genuinely
*fixed*, non-drifting root and, because a Cauchy distribution has much
fatter tails than a normal, produces a visibly wider interval even at the
same nominal level.

## Every episode at once

For a `datestamp()` result with more than one episode, `root_ci_datestamp()`
runs `explosive_root()`/`root_ci()` on each one without a manual loop:

```{r root-ci-datestamp}
root_ci_datestamp(r, ds)
```

Root inference on a very short episode is close to meaningless (too few
points to estimate an AR(1) coefficient precisely) -- filter with
`datestamp(..., min_duration = ...)` before piping in, rather than expecting
`root_ci_datestamp()` to second-guess what counts as "too short."
