---
title: "Root Inference: How Fast Is the Bubble Growing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Root Inference: How Fast Is the Bubble Growing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, warning = FALSE, message = FALSE,
  comment = "#>"
)
```

```{r load}
library(exuber)
```

## A different question from "is there a bubble"

`radf()`, `datestamp()` and the `dating_*()`, `_test()`, `monitor()` and `monitor_*()` families all answer some version of the question of whether there is a bubble and when it happened. None of them says anything about its magnitude. Once an explosive episode is dated, how fast is the underlying autoregressive root growing? `rootstamp()` (Phillips & Magdalinos 2007; Guo, Sun & Wang 2019) answers that question. It is a follow-up step after detection and dating, and it replaces neither.

The function fits a no-intercept AR(1), `y_t = rho * y_{t-1} + e_t`, over a given sub-sample. It reports the estimate of `rho` with a confidence interval and the implied doubling time, `log(2) / log(rho)`, which is the number of periods the bubble needs to double in size at the estimated growth rate. There are two methods for two starting points. The default method takes a numeric sub-sample and fits it once, with one confidence interval. The `radf_obj` method takes a `radf_obj` together with its `datestamp()` result and fits every episode at once, with no manual loop. Neither method returns the `radf_obj` class, so `rootstamp()` does not work with `summary()`, `tidy()` or `autoplot()` (see `vignette("naming-and-analysis")`).

## Detect, date, then estimate the root

The series has a unit-root run-up followed by an explosive regime with `rho = 1.04`:

```{r dgp}
y <- sim_psy1(n = 100, te = 60, tf = 100, c = 0.04, alpha = 0, sigma = 1, seed = 2026)
```

We first detect and date the episode in the usual way:

```{r detect}
r <- radf(y, minw = 20)
cv <- radf_mc_cv(length(y), minw = 20, nrep = 300, seed = 4)
ds <- datestamp(r, cv = cv, min_duration = 3)
ds
```

Then we estimate the root over the detected episode. The default method takes the sub-sample directly, which we slice with the `Start` and `End` of the episode:

```{r root}
ep <- ds[["series1"]]
rootstamp(y[ep$Start[1]:ep$End[1]]) # normal-t interval (Guo, Sun & Wang 2019), true rho = 1.04
rootstamp(y[ep$Start[1]:ep$End[1]], type = "cauchy") # fixed-root Cauchy interval (Phillips & Magdalinos 2007)
```

The estimate of `rho` is close to the true value of 1.04. The output also reports `rho_ci` and the implied `doubling_time` and `doubling_time_ci`. The two interval types answer slightly different questions. `type = "normal"`, the default, is the safer choice under drift or weak dependence, and it gives a noticeably tighter interval here. `type = "cauchy"` assumes a fixed root that does not drift. A Cauchy distribution has much fatter tails than a normal one, so this interval is visibly wider even at the same nominal level.

## Every episode at once

When a `datestamp()` result contains more than one episode, the `radf_obj` method runs the default method on each of them without a manual loop. Pass the original `radf()` result and the `datestamp()` result together:

```{r rootstamp-radf-obj}
rootstamp(r, ds)
```

Root inference on a very short episode is close to meaningless, because there are too few points to estimate an AR(1) coefficient precisely. Filter the episodes with `datestamp(..., min_duration = ...)` before passing them in. The method does not decide for you what counts as too short.
