---
title: "Simulation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, warning = FALSE, message = FALSE,
  comment = "#>", fig.width = 8, fig.height = 4
)
```

```{r load}
library(exuber)
library(ggplot2)
```

## What the `sim_*()` family is for

Every test in the package is calibrated and validated on simulated data,
and the same data-generating processes are exported so you can do the
same: check size under a null you choose, check power against a bubble
you choose, or just see what a given test does to a series whose true
break dates you know. There are three kinds:

| Kind | Functions | Returns |
|---|---|---|
| Bubble processes | `sim_psy1()`, `sim_psy2()`, `sim_ps1()`, `sim_blan()`, `sim_evans()`, `sim_div()`, `sim_tree()`, `sim_mar()`, `sim_msbubble()`, `sim_falsebubble()` | One series (`class "sim"`, plots with `autoplot()`) |
| Innovation generators | `sim_innov()`, `sim_vol_garch()`, `sim_vol_break()`, `sim_vol_cir()`, `sim_vol_sv()`, `sim_fi()` | One series of shocks, to feed into a bubble process through its `e` argument |
| Multi-series processes | `sim_common()`, `sim_coexplosive()` | A `data.frame`, one column per series |

Plus `sim_data`/`sim_data_wdate`, a bundled `data.frame` of five of the
classic series (with and without a date column) used throughout the
examples.

Every function takes a `seed` argument, so a series is reproducible
without a surrounding `set.seed()`, and the bubble processes take the
break dates (`te`, `tf`, ...) as arguments, so the truth you are testing
against is explicit.

The bubble processes split into two families. `sim_psy1()`/`sim_psy2()`/
`sim_ps1()` are the *regime* DGPs of Phillips, Shi & Yu (2015) and Phillips
& Shi (2018): a unit root, then an explosive AR(1) between fixed dates,
then a collapse and a return to a unit root -- the design every paper in
this literature uses for size and power. `sim_blan()`, `sim_evans()`,
`sim_tree()`, `sim_mar()` and `sim_msbubble()` are *rational bubble* models:
the explosive behaviour and its collapse are stochastic, driven by a
probability of bursting each period rather than by fixed dates.
`sim_evans()`'s periodically collapsing bubble is the one PSY's GSADF test
was designed to catch, and the one the earlier tests it replaced could
not.

## The PSY experiment

PSY validate the GSADF test on a price built from Lucas-model fundamentals
plus an Evans (1991) bubble. `sim_div()` gives the fundamental price from a
random walk with drift in dividends (West 1988's S&P 500 parameterisation
by default), `sim_evans()` gives the bubble term, and a scaling factor
`kappa` sets how much of the price the bubble accounts for:

```{r psy-experiment}
n <- 200
pf <- sim_div(n, seed = 1)          # fundamental price
pb <- sim_evans(n, seed = 3)        # periodically collapsing bubble
p <- pf + 20 * pb                   # kappa = 20
```

```{r psy-plot}
data.frame(index = seq_len(n), fundamental = pf, price = p) %>%
  tidyr::pivot_longer(-index) %>%
  ggplot(aes(index, value, color = name)) +
  geom_line() +
  labs(y = NULL, x = NULL, color = NULL) +
  theme_exuber()
```

The point of the exercise is that `radf()` finds the episodes:

```{r psy-radf}
cv <- radf_mc_cv(n, seed = 1)
datestamp(radf(p), cv)
```

## Plotting a simulated series

Each series has an `autoplot()` method:

```{r autoplot-sim, fig.height = 3}
sim_psy2(100, seed = 1) %>%
  autoplot()
```

Several at once go into a `data.frame` -- which is also exactly what
`radf()` takes, so the same object serves both:

```{r sims}
sims <- data.frame(
  psy1 = sim_psy1(100, seed = 1),
  psy2 = sim_psy2(100, seed = 2),
  evans = sim_evans(100, seed = 3),
  blan = sim_blan(100, seed = 4)
)
```

```{r sims-plot, fig.height = 5}
sims %>%
  dplyr::mutate(dplyr::across(dplyr::everything(), as.numeric), index = dplyr::row_number()) %>%
  tidyr::pivot_longer(-index, names_to = "id") %>%
  ggplot(aes(index, value)) +
  geom_line() +
  facet_wrap(~id, scales = "free_y") +
  theme_exuber()
```

## Swapping the innovations

By default the regime DGPs are driven by i.i.d. Gaussian shocks. Every
bubble process accepts a vector of innovations through `e` instead, and
the innovation generators exist to fill it: heavy-tailed or skewed
marginals (`sim_innov()`), conditional heteroskedasticity
(`sim_vol_garch()`), a one-off variance break (`sim_vol_break()`),
stochastic volatility (`sim_vol_cir()`, `sim_vol_sv()`) or long memory
(`sim_fi()`). The bubble dates stay where you put them; only the noise
changes -- which is how the volatility-robust tests in
`vignette("volatility-robust-radf")` are compared against plain `radf()`
on an equal footing:

```{r innovations, fig.height = 3}
# Same bubble, same seed, volatility tripling half-way through the sample
sim_psy1(n = 200, seed = 1, e = sim_vol_break(199, seed = 1)) %>%
  autoplot()
```

The innovation vector is one shorter than `n` because the first
observation is the starting value, not a shock.

## Multi-series processes

`sim_common()` draws `n_series` series that share one latent bubble plus
idiosyncratic noise, the design behind the panel test `radf_common()`;
`sim_coexplosive()` draws a pair where `y` is a linear function of a
(possibly lagged) explosive `x`, the design behind `cobubble_test()` and
`contagion_reg()` (see `vignette("co-explosivity")`):

```{r multi}
head(sim_common(n_series = 3, n = 100, seed = 1))
head(sim_coexplosive(n = 100, lag = 2, seed = 1))
```

## Which to reach for

- Size or power of a test against a bubble with **known dates**:
  `sim_psy1()` (one episode), `sim_psy2()` (two), `sim_ps1()` (one, with a
  distinct collapse regime -- the design the `dating_*()` estimators
  assume).
- A **rational**, stochastically collapsing bubble: `sim_evans()` on its
  own or on top of `sim_div()` fundamentals, as in PSY; `sim_blan()` for
  the Blanchard version; `sim_tree()`, `sim_mar()`, `sim_msbubble()` for
  the newer ones.
- A **null that looks like a bubble** but isn't: `sim_falsebubble()`.
- Non-Gaussian or **heteroskedastic** shocks under any of the above: build
  the shocks with an innovation generator and pass them as `e`.
- **Several series**: `sim_common()` for a shared bubble,
  `sim_coexplosive()` for a linked pair, or a `data.frame` of independent
  draws.
