---
title: "Simulation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The exuber package has numerous functions which allow the creation of time series which incorporate rational bubble models. These include `sim_dgp1()`, which generates a time series with a single episode of explosive behaviour, `sim_dgp2()` which generates a time series with two episodes of explosive behaviour. Two rational bubble models are also currently included; `sim_blan()` which generates a time series containing rational bubbles as proposed by Blanchard (1979). The final function, `sim_evans()` generates a periodically collapsing rational bubble model series. It is the latter type of bubbles that the GSADF test that PSY attempts to detect.

To test the GSADF test and its associated date-stamping strategy, the BSADF test, PSY utilise a model of dividends and incorporate an Evans (1991) type bubble. We replicate that here to show 1) the ability of exuber to generate realistic simulated time series which incorporate bubbles and 2) to demonstrate the capabilities of `radf()` to detect them.

```{r load, echo=FALSE, message=FALSE}
library(exuber)
library(ggplot2)
library(dplyr)
```

Let's start be generating a realistic prices series incorporating an Evans (1991) type periodically collapsing bubble. We will use the monthly parameterisation detailed in PSY (2015a), which corresponds to the empirical values from the S&P 500 (see later example). exuber contains a function `sim_div()` which can be used to generate simulated dividends streams from a random walk with drift model.

Need to make all of these clearer and more consistent with either Evans or Phillips (i.e., need to make a decision)

```{r generate simulated data}
set.seed(125)
# The fundamental value from the Lucas pricing model
pf <- sim_div(400)

# The Evans bubble term
pb <- sim_evans(400)

# the scaling factor for the bubble
kappa <- 20

# The simulated price
p <- pf + kappa*pb
```

We can now plot this data to see what it looks like:

```{r plot simulation, eval = TRUE, fig.width = 9}
tibble(
  index = 1:NROW(p),
  "Price" = p) %>% 
  ggplot(aes(x = index, y = Price)) + 
  geom_line() +
  theme_bw() +
  labs(y = "", x = "")
```

As we can see it looks very typical for the kind of data we wish to analyse with some peaks and troughs which may be indicative of explosive sub-periods. Let us conduct a more formal analysis:

```{r test simulated data, eval = TRUE}
test <- radf(p)
summary(test)
```

# Simulated and MC
