---
title: "Volatility-Robust Alternatives to radf()"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Volatility-Robust Alternatives to radf()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, warning = FALSE, message = FALSE,
  comment = "#>"
)
```

```{r load}
library(exuber)
```

## The shared problem

Plain `radf()` assumes constant innovation variance. Real series rarely
have that, and under time-varying volatility its standard critical values
no longer control size. exuber has several fixes for this, each taking a
structurally different approach. Four of the five below (`radf_sign()`,
`radf_sign_dm()`, `radf_kp()`, `radf_sbz()`) keep the `radf_obj` class and
now have full `summary()`/`datestamp()`/`tidy()`/`autoplot()` support, the
same as plain `radf()` (see `vignette("naming-and-analysis")` for exactly
how each one plugs into the pipeline and how that was validated);
`radf_sbz_union()` doesn't -- it bundles `radf_sbz()`'s statistic and the
classic `supDF` into one union-of-rejections call with its own class, not
a `radf_obj`, so none of those four generics apply to it, only its own
`print()`/`autoplot()`. `radf_tt()`'s time-deformation approach has its
own dedicated vignette, `vignette("radf-tt")`; this one covers the rest.

| Function | Paper | Approach |
|---|---|---|
| `radf_sign()` / `radf_sign_dm()` | Harvey, Leybourne & Zu (2020) | Transform to the *cumulated sign* of first differences -- exactly invariant to any heteroskedasticity pattern, no bootstrap needed. `_dm` demeans first for level-shift robustness (Harvey, Leybourne, Tatlow & Zu 2025). |
| `radf_kp()` | Harvey, Leybourne, Taylor & Zu (2024) | *Purge* volatility: divide each first difference by a kernel spot-volatility estimate, cumulate, then run ordinary PSY on the purged series -- null distribution is identical to the standard homoskedastic one. |
| `radf_sbz()` / `radf_sbz_cv()` | Harvey, Leybourne & Zu (2019) | A WLS-weighted recursive-DF statistic (`supBZ`), *weighting* rather than purging or transforming: same kernel spot-volatility estimator as `radf_kp()`, used as regression weights instead. |
| `radf_sbz_union()` | Harvey, Leybourne & Zu (2019) | `supDF` (classic) and `supBZ` (`radf_sbz()`'s), unioned via a jointly-sized wild bootstrap -- catches whichever of the two has power on a given series. |

## Sign-based: `radf_sign()`

As of 2026-08-18, this also gets full pipeline support -- `radf_sign_cv()`
computes the time-varying `badf_cv`/`bsadf_cv` boundary now, not just the
scalar critical values `summary()` needs (see
`vignette("naming-and-analysis")` for the validation):

```{r sign}
res <- radf_sign(sim_data, minw = 20)
cv <- radf_sign_cv(n = 100, minw = 20)
summary(res, cv = cv)
datestamp(res, cv = cv)
```

`psy2` and `evans` clear their 99% critical values comfortably; the others
don't on this panel -- a realistic mixed result, not every series in a
demo panel is meant to look explosive.

## Kernel-purged: `radf_kp()`

Because `radf_kp()` purges volatility and then calls `radf()` unmodified,
it gets full pipeline support too, the simplest way of all (no new
critical-value machinery at all -- `radf_mc_cv()` applies unmodified):

```{r kp}
res_kp <- radf_kp(sim_data, minw = 20)
cv_kp <- radf_mc_cv(n = attr(res_kp, "n"), minw = 20)
summary(res_kp, cv = cv_kp)
```

## WLS + kernel volatility: `radf_sbz()`

As of 2026-08-22, `radf_sbz()` is split from the union test into its own
statistic function, with `radf_sbz_cv()` computing the time-varying
`badf_cv`/`bsadf_cv` boundary the same way `radf_tt_cv()`/`radf_sign_cv()`
do (see `vignette("naming-and-analysis")` for the validation), so it gets
full pipeline support:

```{r sbz}
res_sbz <- radf_sbz(sim_data, minw = 20)
cv_sbz <- radf_sbz_cv(sim_data, minw = 20, nboot = 200, seed = 1)
summary(res_sbz, cv = cv_sbz)
```

The kernel-volatility weighting that makes `supBZ` heteroskedasticity-robust
also trades away some power relative to plain `radf()` on series whose
bubbles aren't accompanied by a volatility shift -- none of this panel
clears `supBZ`'s 95% critical value here, even though several clear plain
`radf()`'s. That's the same trade-off `radf_sbz_union()` below exists to
hedge against by unioning `supBZ` with the classic `supDF`.

## Union-of-rejections: `radf_sbz_union()`

```{r sbz-union}
radf_sbz_union(sim_data, nboot = 200, seed = 1)
```

`supDF` is the classic PWY statistic, `supBZ` the WLS-weighted version (the
same one `radf_sbz()` now returns on its own), and `U` their union -- each
with its own bootstrap p-value, so a series can be flagged by one without
the other: `psy2` is clearly significant across all three here, while the
rest of this panel sit closer to the boundary, with `supBZ`'s p-values
consistently the least significant of the three on this draw. `U`'s value
is *defined* using a bootstrap-derived `supDF`/`supBZ` scaling ratio, and
its size guarantee requires `supDF`/`supBZ` bootstrap draws paired from the
same resampled series per replicate -- both reasons `radf_sbz_union()`
can't be reconstructed from separately calling `radf_sbz_cv()` and plain
`radf_wb_cv()`, and why it stays a single bundled call with its own class
rather than a `radf_obj`.

## Which to reach for

- Want exact invariance to *any* heteroskedasticity pattern with no
  bootstrap at all: `radf_sign()` (or `radf_sign_dm()` if a level shift,
  not just volatility, is a concern). Also full pipeline support.
- Want to stay closest to plain `radf()`, no new critical-value machinery
  at all: `radf_kp()`.
- Want the WLS efficiency gain with full `datestamp()`/`autoplot()`
  support: `radf_sbz()` + `radf_sbz_cv()`.
- Want to hedge between the classic and WLS-weighted statistics on the
  same series, and don't need `datestamp()`/`autoplot()`:
  `radf_sbz_union()` -- the one function in this group with no pipeline
  support, since `U` bundles both statistics and their (scalar-only)
  joint critical value in one call rather than returning a `radf_obj`.
- Volatility is the whole story and a bootstrap-free, time-deformation
  approach is preferred: `radf_tt()`, see `vignette("radf-tt")`.
- Volatility is genuinely unknown/complex and a bootstrap is acceptable:
  plain `radf_wb_cv()` remains the general-purpose choice.
