---
title: "Volatility-Robust Alternatives to radf()"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Volatility-Robust Alternatives to radf()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, warning = FALSE, message = FALSE,
  comment = "#>"
)
```

```{r load}
library(exuber)
```

## The shared problem

Plain `radf()` assumes constant innovation variance. Real series rarely
have that, and under time-varying volatility its standard critical values
no longer control size. exuber has several fixes for this, each taking a
structurally different approach, and -- unlike most of the other new
functions -- all four here keep the `radf_obj` class, so `summary()`/`tidy()`
still work the same way they do for plain `radf()` (see
`vignette("naming-and-analysis")` for exactly how far each one plugs into
the pipeline). `radf_tt()`'s time-deformation approach has its own
dedicated vignette, `vignette("radf-tt")`; this one covers the rest.

| Function | Paper | Approach |
|---|---|---|
| `radf_sign()` / `radf_sign_dm()` | Harvey, Leybourne & Zu (2020) | Transform to the *cumulated sign* of first differences -- exactly invariant to any heteroskedasticity pattern, no bootstrap needed. `_dm` demeans first for level-shift robustness (Harvey, Leybourne, Tatlow & Zu 2025). |
| `radf_kp()` | Harvey, Leybourne, Taylor & Zu (2024) | *Purge* volatility: divide each first difference by a kernel spot-volatility estimate, cumulate, then run ordinary PSY on the purged series -- null distribution is identical to the standard homoskedastic one. |
| `radf_sbz_cv()` | Harvey, Leybourne & Zu (2019) | A WLS-weighted test using the same kernel volatility estimator, unioned with the classic supDF via a jointly-sized wild bootstrap. |

## Sign-based: `radf_sign()`

```{r sign}
res <- radf_sign(sim_data, minw = 20)
cv <- radf_sign_cv(n = 100, minw = 20)
summary(res, cv = cv)
```

`psy2` and `evans` clear their 99% critical values comfortably; the others
don't on this panel -- a realistic mixed result, not every series in a
demo panel is meant to look explosive.

## Kernel-purged: `radf_kp()`

Because `radf_kp()` purges volatility and then calls `radf()` unmodified,
it gets **full** pipeline support (`summary()`, `datestamp()`, `tidy()`,
`autoplot()` all work, unlike the sign-based and SBZ variants):

```{r kp}
res_kp <- radf_kp(sim_data, minw = 20)
cv_kp <- radf_mc_cv(n = attr(res_kp, "n"), minw = 20)
summary(res_kp, cv = cv_kp)
```

## WLS + kernel volatility with union-of-rejections: `radf_sbz_cv()`

```{r sbz}
radf_sbz_cv(sim_data, nboot = 200, seed = 1)
```

`supDF` is the classic PWY statistic, `supBZ` the WLS-weighted version, and
`U` their union -- each with its own bootstrap p-value, so a series can be
flagged by one without the other: `psy2` is clearly significant across all
three here, while the rest of this panel sit closer to the boundary, with
`supBZ`'s p-values consistently the least significant of the three on this
draw.

## Which to reach for

- Want exact invariance to *any* heteroskedasticity pattern with no
  bootstrap at all: `radf_sign()` (or `radf_sign_dm()` if a level shift,
  not just volatility, is a concern).
- Want to stay inside the full `summary()`/`datestamp()`/`tidy()`/
  `autoplot()` pipeline: `radf_kp()` -- the only one of this group with no
  gaps there.
- Want the WLS efficiency gain and a union test against the classic
  statistic: `radf_sbz_cv()`.
- Volatility is the whole story and a bootstrap-free, time-deformation
  approach is preferred: `radf_tt()`, see `vignette("radf-tt")`.
- Volatility is genuinely unknown/complex and a bootstrap is acceptable:
  plain `radf_wb_cv()` remains the general-purpose choice.
